---
title: "ACAT (20 genes)"
author: "XSun"
date: "2023-01-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
suppressMessages(library(data.table))
suppressMessages(library(htmltools))
```

# Results (calibrated ACAT p-values)

For each trait (trait X), use the random ACAT p-values related to trait X (all factor-trait X pairs) as null distribution, then correct the real ACAT p-values. All calibration will be done within each trait.

We calibrated p-values according to: 

p_corr[t] <- sum(p_real[t] > p_random)/length(p_random)

For those p_corr == 0, we set p_corr[p_corr == 0] <- 0.5/length(p_random)

# ACAT summary

The following tables show the top pairs with lowest calibrated ACAT p-values for each trait, only pairs with >=3 supporting SNPs are included.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE,fig.height=10, fig.width=15}
traits <- c("EUR.UC","EUR.CD","EUR.IBD","mch","mchc","mcv","rdw","ret","baso","plt","pct","pdw","mpv","hct","hgb","mono","eo","gran","lymph","myeloid_wbc","neut","rbc","wbc","ukb.allasthma","allergy","T1D")
path <- c("4.B_cell_kegg","6.CD14_positive_monocyte_kegg","8.CD15_positive_leukocyte_kegg","16.platelet_kegg","10.T_cell_kegg","14.thymocyte_kegg")


#load("/project2/xinhe/xsun/CEDAR/ACAT_randomsnp_FDR_corrected/results/acat_real.rdata")
load("/project2/xinhe/xsun/CEDAR/ACAT_randomsnp_FDR_corrected/results/calibratedbytrait_acat_real_p.rdata")
pathway <- read.table("/project2/xinhe/xsun/pathways/kegg/pathway_name.txt",stringsAsFactors = F, fill = T, sep = "\t",colClasses = "character")
colnames(pathway) <- c("code","name")
#for (i in 1:length(acat_allcelltypes)){
  i <- 1
  acat_celltype <- acat_corr_all[[i]]
  
  file_snp_num <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/10.20genes_rerun/2.ECT/results/number_snp_fdr02_allpairs.rdata")
  load(file_snp_num)
  
  file_expr <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/1.find_factors/all.pwy_g_expr_.rdata")
  load(file_expr)
  
  acat_summary <- c()
  acat_summary_list <- list()
  for (j in 1:length(snp_num_fdr02)) {
    
    trait <- names(snp_num_fdr02)[j]
    index_acat <- grep(pattern = trait, names(acat_celltype))
    acat_trait <- acat_celltype[index_acat]
    
    snp_trait <- snp_num_fdr02[[j]]
    snp_trait <- snp_trait[snp_trait >=3]
    snp_trait <- data.frame(cbind(paste0(trait,"_",names(snp_trait)),snp_trait))
    colnames(snp_trait) <- c("pair","num_snp")
    
    acat_trait <- data.frame(cbind(names(acat_trait),acat_trait))
    colnames(acat_trait) <- c("pair","acat_p")
    acat_trait_snp <- merge(acat_trait,snp_trait,by = "pair")

    acat_trait_snp_sort <- acat_trait_snp[order(acat_trait_snp$acat_p,decreasing = F),]
    acat_trait_snp_sort_10 <- acat_trait_snp_sort[1:10,]
    
    pwy_name <- c()
    for (t in 1:nrow(acat_trait_snp_sort_10)) {
      
      spl <- unlist(strsplit(acat_trait_snp_sort_10$pair[t],split ="_"))
      
      if(length(spl) ==3) {
        pwy <- as.numeric(unlist(strsplit(spl[2],split = "pwy"))[2])
      }else {
        pwy <- as.numeric(unlist(strsplit(spl[3],split = "pwy"))[2])
      }
          
      pwy_code_pair <- as.character(geneexpr_all_pway[[pwy]]$pathway[1])
      pwy_name_pair <- pathway[pathway$code == pwy_code_pair,2]

      pwy_name <- c(pwy_name,pwy_name_pair)
      
    }
    acat_trait_snp_sort_10$pwy_name <- pwy_name
    acat_trait_snp_sort_10 <- as.matrix(acat_trait_snp_sort_10)
    
    acat_summary <- rbind(acat_summary,acat_trait_snp_sort_10)
    acat_summary_list[[j]] <- acat_trait_snp_sort_10
  }
  acat_summary <- data.frame(acat_summary)
  rownames(acat_summary) <- seq(1,nrow(acat_summary),by=1)
  print(names(acat_corr_all)[i])
  colnames(acat_summary)[2] <- "corrected_acat_p"
  DT::datatable(acat_summary)
  #print(acat_summary)

  i <- 2
  acat_celltype <- acat_corr_all[[i]]
  
  file_snp_num <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/10.20genes_rerun/2.ECT/results/number_snp_fdr02_allpairs.rdata")
  load(file_snp_num)
  
  file_expr <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/1.find_factors/all.pwy_g_expr_.rdata")
  load(file_expr)
  
  acat_summary <- c()
  acat_summary_list <- list()
  for (j in 1:length(snp_num_fdr02)) {
    
    trait <- names(snp_num_fdr02)[j]
    index_acat <- grep(pattern = trait, names(acat_celltype))
    acat_trait <- acat_celltype[index_acat]
    
    snp_trait <- snp_num_fdr02[[j]]
    snp_trait <- snp_trait[snp_trait >=3]
    snp_trait <- data.frame(cbind(paste0(trait,"_",names(snp_trait)),snp_trait))
    colnames(snp_trait) <- c("pair","num_snp")
    
    acat_trait <- data.frame(cbind(names(acat_trait),acat_trait))
    colnames(acat_trait) <- c("pair","acat_p")
    acat_trait_snp <- merge(acat_trait,snp_trait,by = "pair")

    acat_trait_snp_sort <- acat_trait_snp[order(acat_trait_snp$acat_p,decreasing = F),]
    acat_trait_snp_sort_10 <- acat_trait_snp_sort[1:10,]
    
    pwy_name <- c()
    for (t in 1:nrow(acat_trait_snp_sort_10)) {
      
      spl <- unlist(strsplit(acat_trait_snp_sort_10$pair[t],split ="_"))
      
      if(length(spl) ==3) {
        pwy <- as.numeric(unlist(strsplit(spl[2],split = "pwy"))[2])
      }else {
        pwy <- as.numeric(unlist(strsplit(spl[3],split = "pwy"))[2])
      }
          
      pwy_code_pair <- as.character(geneexpr_all_pway[[pwy]]$pathway[1])
      pwy_name_pair <- pathway[pathway$code == pwy_code_pair,2]

      pwy_name <- c(pwy_name,pwy_name_pair)
      
    }
    acat_trait_snp_sort_10$pwy_name <- pwy_name
    acat_trait_snp_sort_10 <- as.matrix(acat_trait_snp_sort_10)
    
    acat_summary <- rbind(acat_summary,acat_trait_snp_sort_10)
    acat_summary_list[[j]] <- acat_trait_snp_sort_10
  }
  acat_summary <- data.frame(acat_summary)
  rownames(acat_summary) <- seq(1,nrow(acat_summary),by=1)
  print(names(acat_corr_all)[i])
  colnames(acat_summary)[2] <- "corrected_acat_p"
  DT::datatable(acat_summary)
  
    i <- 3
  acat_celltype <- acat_corr_all[[i]]
  
  file_snp_num <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/10.20genes_rerun/2.ECT/results/number_snp_fdr02_allpairs.rdata")
  load(file_snp_num)
  
  file_expr <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/1.find_factors/all.pwy_g_expr_.rdata")
  load(file_expr)
  
  acat_summary <- c()
  acat_summary_list <- list()
  for (j in 1:length(snp_num_fdr02)) {
    
    trait <- names(snp_num_fdr02)[j]
    index_acat <- grep(pattern = trait, names(acat_celltype))
    acat_trait <- acat_celltype[index_acat]
    
    snp_trait <- snp_num_fdr02[[j]]
    snp_trait <- snp_trait[snp_trait >=3]
    snp_trait <- data.frame(cbind(paste0(trait,"_",names(snp_trait)),snp_trait))
    colnames(snp_trait) <- c("pair","num_snp")
    
    acat_trait <- data.frame(cbind(names(acat_trait),acat_trait))
    colnames(acat_trait) <- c("pair","acat_p")
    acat_trait_snp <- merge(acat_trait,snp_trait,by = "pair")

    acat_trait_snp_sort <- acat_trait_snp[order(acat_trait_snp$acat_p,decreasing = F),]
    acat_trait_snp_sort_10 <- acat_trait_snp_sort[1:10,]
    
    pwy_name <- c()
    for (t in 1:nrow(acat_trait_snp_sort_10)) {
      
      spl <- unlist(strsplit(acat_trait_snp_sort_10$pair[t],split ="_"))
      
      if(length(spl) ==3) {
        pwy <- as.numeric(unlist(strsplit(spl[2],split = "pwy"))[2])
      }else {
        pwy <- as.numeric(unlist(strsplit(spl[3],split = "pwy"))[2])
      }
          
      pwy_code_pair <- as.character(geneexpr_all_pway[[pwy]]$pathway[1])
      pwy_name_pair <- pathway[pathway$code == pwy_code_pair,2]

      pwy_name <- c(pwy_name,pwy_name_pair)
      
    }
    acat_trait_snp_sort_10$pwy_name <- pwy_name
    acat_trait_snp_sort_10 <- as.matrix(acat_trait_snp_sort_10)
    
    acat_summary <- rbind(acat_summary,acat_trait_snp_sort_10)
    acat_summary_list[[j]] <- acat_trait_snp_sort_10
  }
  acat_summary <- data.frame(acat_summary)
  rownames(acat_summary) <- seq(1,nrow(acat_summary),by=1)
  print(names(acat_corr_all)[i])
  colnames(acat_summary)[2] <- "corrected_acat_p"
  DT::datatable(acat_summary)

    i <- 4
  acat_celltype <- acat_corr_all[[i]]
  
  file_snp_num <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/10.20genes_rerun/2.ECT/results/number_snp_fdr02_allpairs.rdata")
  load(file_snp_num)
  
  file_expr <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/1.find_factors/all.pwy_g_expr_.rdata")
  load(file_expr)
  
  acat_summary <- c()
  acat_summary_list <- list()
  for (j in 1:length(snp_num_fdr02)) {
    
    trait <- names(snp_num_fdr02)[j]
    index_acat <- grep(pattern = trait, names(acat_celltype))
    acat_trait <- acat_celltype[index_acat]
    
    snp_trait <- snp_num_fdr02[[j]]
    snp_trait <- snp_trait[snp_trait >=3]
    snp_trait <- data.frame(cbind(paste0(trait,"_",names(snp_trait)),snp_trait))
    colnames(snp_trait) <- c("pair","num_snp")
    
    acat_trait <- data.frame(cbind(names(acat_trait),acat_trait))
    colnames(acat_trait) <- c("pair","acat_p")
    acat_trait_snp <- merge(acat_trait,snp_trait,by = "pair")

    acat_trait_snp_sort <- acat_trait_snp[order(acat_trait_snp$acat_p,decreasing = F),]
    acat_trait_snp_sort_10 <- acat_trait_snp_sort[1:10,]
    
    pwy_name <- c()
    for (t in 1:nrow(acat_trait_snp_sort_10)) {
      
      spl <- unlist(strsplit(acat_trait_snp_sort_10$pair[t],split ="_"))
      
      if(length(spl) ==3) {
        pwy <- as.numeric(unlist(strsplit(spl[2],split = "pwy"))[2])
      }else {
        pwy <- as.numeric(unlist(strsplit(spl[3],split = "pwy"))[2])
      }
          
      pwy_code_pair <- as.character(geneexpr_all_pway[[pwy]]$pathway[1])
      pwy_name_pair <- pathway[pathway$code == pwy_code_pair,2]

      pwy_name <- c(pwy_name,pwy_name_pair)
      
    }
    acat_trait_snp_sort_10$pwy_name <- pwy_name
    acat_trait_snp_sort_10 <- as.matrix(acat_trait_snp_sort_10)
    
    acat_summary <- rbind(acat_summary,acat_trait_snp_sort_10)
    acat_summary_list[[j]] <- acat_trait_snp_sort_10
  }
  acat_summary <- data.frame(acat_summary)
  rownames(acat_summary) <- seq(1,nrow(acat_summary),by=1)
  print(names(acat_corr_all)[i])
  colnames(acat_summary)[2] <- "corrected_acat_p"
  DT::datatable(acat_summary)
  
  
    i <- 5
  acat_celltype <- acat_corr_all[[i]]
  
  file_snp_num <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/10.20genes_rerun/2.ECT/results/number_snp_fdr02_allpairs.rdata")
  load(file_snp_num)
  
  file_expr <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/1.find_factors/all.pwy_g_expr_.rdata")
  load(file_expr)
  
  acat_summary <- c()
  acat_summary_list <- list()
  for (j in 1:length(snp_num_fdr02)) {
    
    trait <- names(snp_num_fdr02)[j]
    index_acat <- grep(pattern = trait, names(acat_celltype))
    acat_trait <- acat_celltype[index_acat]
    
    snp_trait <- snp_num_fdr02[[j]]
    snp_trait <- snp_trait[snp_trait >=3]
    snp_trait <- data.frame(cbind(paste0(trait,"_",names(snp_trait)),snp_trait))
    colnames(snp_trait) <- c("pair","num_snp")
    
    acat_trait <- data.frame(cbind(names(acat_trait),acat_trait))
    colnames(acat_trait) <- c("pair","acat_p")
    acat_trait_snp <- merge(acat_trait,snp_trait,by = "pair")

    acat_trait_snp_sort <- acat_trait_snp[order(acat_trait_snp$acat_p,decreasing = F),]
    acat_trait_snp_sort_10 <- acat_trait_snp_sort[1:10,]
    
    pwy_name <- c()
    for (t in 1:nrow(acat_trait_snp_sort_10)) {
      
      spl <- unlist(strsplit(acat_trait_snp_sort_10$pair[t],split ="_"))
      
      if(length(spl) ==3) {
        pwy <- as.numeric(unlist(strsplit(spl[2],split = "pwy"))[2])
      }else {
        pwy <- as.numeric(unlist(strsplit(spl[3],split = "pwy"))[2])
      }
          
      pwy_code_pair <- as.character(geneexpr_all_pway[[pwy]]$pathway[1])
      pwy_name_pair <- pathway[pathway$code == pwy_code_pair,2]

      pwy_name <- c(pwy_name,pwy_name_pair)
      
    }
    acat_trait_snp_sort_10$pwy_name <- pwy_name
    acat_trait_snp_sort_10 <- as.matrix(acat_trait_snp_sort_10)
    
    acat_summary <- rbind(acat_summary,acat_trait_snp_sort_10)
    acat_summary_list[[j]] <- acat_trait_snp_sort_10
  }
  acat_summary <- data.frame(acat_summary)
  rownames(acat_summary) <- seq(1,nrow(acat_summary),by=1)
  print(names(acat_corr_all)[i])
  colnames(acat_summary)[2] <- "corrected_acat_p"
  DT::datatable(acat_summary)
  
    i <- 6
  acat_celltype <- acat_corr_all[[i]]
  
  file_snp_num <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/10.20genes_rerun/2.ECT/results/number_snp_fdr02_allpairs.rdata")
  load(file_snp_num)
  
  file_expr <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/1.find_factors/all.pwy_g_expr_.rdata")
  load(file_expr)
  
  acat_summary <- c()
  acat_summary_list <- list()
  for (j in 1:length(snp_num_fdr02)) {
    
    trait <- names(snp_num_fdr02)[j]
    index_acat <- grep(pattern = trait, names(acat_celltype))
    acat_trait <- acat_celltype[index_acat]
    
    snp_trait <- snp_num_fdr02[[j]]
    snp_trait <- snp_trait[snp_trait >=3]
    snp_trait <- data.frame(cbind(paste0(trait,"_",names(snp_trait)),snp_trait))
    colnames(snp_trait) <- c("pair","num_snp")
    
    acat_trait <- data.frame(cbind(names(acat_trait),acat_trait))
    colnames(acat_trait) <- c("pair","acat_p")
    acat_trait_snp <- merge(acat_trait,snp_trait,by = "pair")

    acat_trait_snp_sort <- acat_trait_snp[order(acat_trait_snp$acat_p,decreasing = F),]
    acat_trait_snp_sort_10 <- acat_trait_snp_sort[1:10,]
    
    pwy_name <- c()
    for (t in 1:nrow(acat_trait_snp_sort_10)) {
      
      spl <- unlist(strsplit(acat_trait_snp_sort_10$pair[t],split ="_"))
      
      if(length(spl) ==3) {
        pwy <- as.numeric(unlist(strsplit(spl[2],split = "pwy"))[2])
      }else {
        pwy <- as.numeric(unlist(strsplit(spl[3],split = "pwy"))[2])
      }
          
      pwy_code_pair <- as.character(geneexpr_all_pway[[pwy]]$pathway[1])
      pwy_name_pair <- pathway[pathway$code == pwy_code_pair,2]

      pwy_name <- c(pwy_name,pwy_name_pair)
      
    }
    acat_trait_snp_sort_10$pwy_name <- pwy_name
    acat_trait_snp_sort_10 <- as.matrix(acat_trait_snp_sort_10)
    
    acat_summary <- rbind(acat_summary,acat_trait_snp_sort_10)
    acat_summary_list[[j]] <- acat_trait_snp_sort_10
  }
  acat_summary <- data.frame(acat_summary)
  rownames(acat_summary) <- seq(1,nrow(acat_summary),by=1)
  print(names(acat_corr_all)[i])
  colnames(acat_summary)[2] <- "corrected_acat_p"
  DT::datatable(acat_summary)
  
#}

```


# Summary plots

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE,fig.height=10, fig.width=18}

##### qqplot
load("/project2/xinhe/xsun/CEDAR/ACAT_randomsnp_FDR_20genes/results/calibratedbytrait_acat_real_p.rdata")
type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")
plot <- list()
for (i in 1:length(type)){
  
  p <- acat_corr_all[[i]]

  plotdata <- data.frame(observed = -log10(sort(p)),
                         expected = -log10(ppoints(length(p))))
  #  
  
  plot[[i]] <- ggplot(plotdata) + theme_bw(base_line_size =0.3) +
    geom_point(aes(expected, observed), shape = 1, size = 1.5,color = "steelblue") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.6, color = "red") +
    ggtitle(type[i]) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = expression(paste("Expected -log"[10],"(p-value)")),
         y = expression(paste("Observed -log"[10],"(p-value)"))) +
    
    theme(axis.title.x = element_text(size = 10),
          axis.text.x = element_text(size = 8, color = "black"),
          axis.title.y = element_text(size = 10),
          axis.text.y = element_text(size = 8, color = "black"),
          text= element_text(family="Times New Roman")) 
  
  
}
all <- grid.arrange(plot[[1]],plot[[2]],plot[[3]],plot[[4]],plot[[5]],plot[[6]], nrow = 2)


for (i in 1:length(type)){
  
  p <- as.data.frame(acat_corr_all[[i]])
  colnames(p) <- "p"
  #  
  
  
  plot[[i]] <- ggplot(p, aes(x=p)) + geom_histogram(breaks = seq(0, 1, by = 0.05),color="white",fill = "steelblue")+ 
    theme_bw(base_line_size =0.3) +
    ggtitle(type[i]) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = "Corrected ACAT p-values",
         y = "Count") +
    
    theme(axis.title.x = element_text(size = 13),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12, color = "black"),
          text= element_text(family="Times")) 
  
  
}
all <- grid.arrange(plot[[1]],plot[[2]],plot[[3]],plot[[4]],plot[[5]],plot[[6]], nrow = 2)
```