---
title: "ECT result analyzing"
author: "XSun"
date: "2024-02-22"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

## Functions used

```{r message=FALSE}
library(tibble)
library(dplyr)
library(ggplot2)

library(tidyverse)
library(ComplexHeatmap)
library(circlize)

traits <- c("eo", "gran", "hct", "hgb", "lymph", "mch", "mchc", "mcv", 
            "mpv", "mono", "myeloid_wbc", "neut", "plt", "pdw", "pct", 
            "rbc", "rdw", "ret", "wbc", "baso", "T1D", "ukb.allasthma", 
            "EUR.IBD", "EUR.CD", "EUR.UC", "allergy")
celltypes <- c("B_cell","CD14_positive_monocyte","CD15_positive_leukocyte","platelet","T_cell","thymocyte")
type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")


plot_selected_pwys <- function(trait, summary_df) {
  
  # Subset summary_df for the given trait
  selected <- summary_df[summary_df$trait == trait, ]
  
  # Handle duplicated pathway names by appending a space
  index <- which(duplicated(selected$pwy_name))
  selected$pwy_name[index] <- paste0(selected$pwy_name[index], " ")
  
  # Transform p-values and order by -log10(pval)
  selected$lgectpval <- -log10(selected$pval_ECT)
  selected <- selected[order(selected$lgectpval, decreasing = TRUE), ]
  
  # Append note to the last num_snp_pass_cutoff value
  selected$num_snp_pass_cutoff[nrow(selected)] <- paste0(selected$num_snp_pass_cutoff[nrow(selected)], " (# of supporting variants)")
  
  # Set pathway names as a factor with levels in the desired order
  selected$pwy_name <- factor(selected$pwy_name, levels = rev(unique(selected$pwy_name)))
  
  # Create the plot
  p <- ggplot(selected) +
    theme_bw(base_line_size = 0.3) +
    geom_col(aes(x = lgectpval, y = pwy_name, fill = celltypes), width = 0.6) +
    geom_text(aes(0, y = pwy_name, label = pwy_name), hjust = 0, nudge_x = 0.1, colour = "white", size = 3, family = "Times") + 
    geom_text(aes(lgectpval, y = pwy_name, label = num_snp_pass_cutoff), hjust = 0, nudge_x = 0.01, colour = "black", size = 3, family = "Times") +
    labs(x = expression(paste("-log"[10], "(ECT p-value)")), y = "Pathways") +
    theme(axis.text.y = element_blank(), text = element_text(family = "Times")) +
    xlim(0, ceiling(max(selected$lgectpval))) +
    ggtitle(trait)
  
  return(p)
}


plot_selected_traits <- function(pwy, summary_df) {
  
  # Subset summary_df for the given trait
  selected <- summary_df[summary_df$pwy_name == pwy, ]
  
  # Handle duplicated trait names by appending a space
  index <- which(duplicated(selected$trait))
  selected$trait[index] <- paste0(selected$trait[index], " ")
  
  # Transform p-values and order by -log10(pval)
  selected$lgectpval <- -log10(selected$pval_ECT)
  selected <- selected[order(selected$lgectpval, decreasing = TRUE), ]
  
  # Append note to the last num_snp_pass_cutoff value
  selected$num_snp_pass_cutoff[nrow(selected)] <- paste0(selected$num_snp_pass_cutoff[nrow(selected)], " (# of supporting variants)")
  
  # Set pathway names as a factor with levels in the desired order
  selected$trait <- factor(selected$trait, levels = rev(unique(selected$trait)))
  
  # Create the plot
  p <- ggplot(selected) +
    theme_bw(base_line_size = 0.3) +
    geom_col(aes(x = lgectpval, y = trait, fill = celltypes), width = 0.6) +
    geom_text(aes(0, y = trait, label = trait), hjust = 0, nudge_x = 0.1, colour = "white", size = 3, family = "Times") + 
    geom_text(aes(lgectpval, y = trait, label = num_snp_pass_cutoff), hjust = 0, nudge_x = 0.01, colour = "black", size = 3, family = "Times") +
    labs(x = expression(paste("-log"[10], "(ECT p-value)")), y = "Traits") +
    theme(axis.text.y = element_blank(), text = element_text(family = "Times")) +
    xlim(0, ceiling(max(selected$lgectpval))) +
    ggtitle(pwy)
  
  return(p)
}

generate_heatmap <- function(df_all) {
  # Count the number of occurrences for each trait-pwy_name pair
  count_data <- df_all %>%
    dplyr::group_by(trait, pwy_name) %>%
    dplyr::summarise(n = n(), .groups = 'drop')
  
  # Create a matrix for the heatmap
  heatmap_data <- count_data %>%
    tidyr::pivot_wider(names_from = pwy_name, values_from = n, values_fill = list(n = 0)) %>%
    column_to_rownames(var = "trait") %>%
    as.matrix()
  
  # Determine the range of integer values
  value_range <- range(heatmap_data, na.rm = TRUE)
  value_range <- c(floor(value_range[1]), ceiling(value_range[2]))  # Ensure they are integers
  
  # Create a sequence of integers within this range
  int_seq <- seq(value_range[1], value_range[2], by = 1)
  colors <- colorRampPalette(c("white", "red"))(length(int_seq))
  
  # Create a named vector where each integer is associated with a specific color
  color_mapping <- setNames(colors, int_seq)
  
  # Generate the heatmap with the specified color mapping
  ComplexHeatmap::Heatmap(heatmap_data, name = "count",
                          column_title = "Pathway Name", row_title = "Trait",
                          col = color_mapping,  # Use the discrete color mapping
                          border = FALSE, 
                          show_row_dend = FALSE, 
                          show_column_dend = FALSE)
}

```

# Pathways & traits heat map 

## ECT FDR < 0.2

```{r fig.width=15, fig.height=10}
load("/project/xinhe/xsun/pathway_factor/analysis/1.ECT/results/ECT_summary_mbeadded_ectfdr02_allcelltypes.rdata")
generate_heatmap(df_all = df_all)

```


## ECT FDR < 0.1

```{r }
load("/project/xinhe/xsun/pathway_factor/analysis/1.ECT/results/ECT_summary_mbeadded_ectfdr02_allcelltypes.rdata")

df_all <- df_all[df_all$fdr_ECT < 0.1,]
generate_heatmap(df_all = df_all)

```

# Pathways identified across Traits at various cutoffs

## ECT FDR < 0.2

```{r}
load("/project/xinhe/xsun/pathway_factor/analysis/1.ECT/results/ECT_summary_mbeadded_ectfdr02_allcelltypes.rdata")

num_pwy_pertrait <- lapply(traits,function(trait) {
       sum(!duplicated(df_all$factor[df_all$trait == trait]))
  })

num_pwy_pertrait <- unlist(num_pwy_pertrait)
names(num_pwy_pertrait) <- traits

data_long <- enframe(num_pwy_pertrait, name = "trait", value = "count")

# Plot
ggplot(data_long, aes(x = trait, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(title = "Number of Pathways per Trait at FDR < 0.2",
       x = "Trait",
       y = "Number of Pathways") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot_selected_pwys(trait = "baso",summary_df = df_all)
plot_selected_pwys(trait = "ret",summary_df = df_all)
plot_selected_pwys(trait = "lymph",summary_df = df_all)


```

## ECT FDR < 0.1

```{r}
load("/project/xinhe/xsun/pathway_factor/analysis/1.ECT/results/ECT_summary_mbeadded_ectfdr02_allcelltypes.rdata")

df_all <- df_all[df_all$fdr_ECT < 0.1,]

num_pwy_pertrait <- lapply(traits,function(trait) {
       sum(!duplicated(df_all$factor[df_all$trait == trait]))
  })

num_pwy_pertrait <- unlist(num_pwy_pertrait)
names(num_pwy_pertrait) <- traits

data_long <- enframe(num_pwy_pertrait, name = "trait", value = "count")

# Plot
ggplot(data_long, aes(x = trait, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(title = "Number of Pathways per Trait at FDR < 0.1",
       x = "Trait",
       y = "Number of Pathways") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot_selected_pwys(trait = "ret",summary_df = df_all)
plot_selected_pwys(trait = "pct",summary_df = df_all)
plot_selected_pwys(trait = "lymph",summary_df = df_all)


```


# Pathways involved in multiplt traits

## ECT FDR < 0.2

```{r}

load("/project/xinhe/xsun/pathway_factor/analysis/1.ECT/results/ECT_summary_mbeadded_ectfdr02_allcelltypes.rdata")

pwys <- df_all$pwy_code[!duplicated(df_all$pwy_code)]

num_trait_perpwy <- lapply(pwys,function(pwy) {
       sum(!duplicated(df_all$trait[df_all$pwy_code == pwy]))
  })

num_trait_perpwy <- unlist(num_trait_perpwy)
names(num_trait_perpwy) <- df_all$pwy_name[match(pwys, df_all$pwy_code)]

num_trait_perpwy <- num_trait_perpwy[num_trait_perpwy>1]

data_long <- enframe(num_trait_perpwy, name = "pwy", value = "count")

# Plot
ggplot(data_long, aes(x = pwy, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(title = "Number of Traits per pwy at FDR < 0.1",
       x = "pwy",
       y = "Number of Traits") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))





plot_selected_traits("Cellular senescence",df_all)
plot_selected_traits("cAMP signaling pathway",df_all)





```





## ECT FDR < 0.1

```{r}

load("/project/xinhe/xsun/pathway_factor/analysis/1.ECT/results/ECT_summary_mbeadded_ectfdr02_allcelltypes.rdata")

df_all <- df_all[df_all$fdr_ECT < 0.1,]

pwys <- df_all$pwy_code[!duplicated(df_all$pwy_code)]

num_trait_perpwy <- lapply(pwys,function(pwy) {
       sum(!duplicated(df_all$trait[df_all$pwy_code == pwy]))
  })

num_trait_perpwy <- unlist(num_trait_perpwy)
names(num_trait_perpwy) <- df_all$pwy_name[match(pwys, df_all$pwy_code)]

num_trait_perpwy <- num_trait_perpwy[num_trait_perpwy>1]

data_long <- enframe(num_trait_perpwy, name = "pwy", value = "count")

# Plot
ggplot(data_long, aes(x = pwy, y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(title = "Number of Traits per pwy at FDR < 0.1",
       x = "pwy",
       y = "Number of Traits") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot_selected_traits("Small cell lung cancer",df_all)
```




# Genes close to supporting variants -- for pairs with ECT FDR < 0.1


```{r}
load("/project/xinhe/xsun/pathway_factor/analysis/3.results_analysis/results/nearby_gene_added.rdata")

df <- df_all[,c("celltypes","factor","trait","pwy_name","fdr_ECT","num_snp_pass_cutoff","gene_near_supp_snps")]
colnames(df) <- c("celltypes","factor","trait","pwy_name","fdr_ECT","#of supporting SNPs","gene near supporting SNPs")
DT::datatable(df)
```


[Locus Zoom plots](https://uchicago.box.com/s/8f349903pupvsubmjvbs6daawrjjla1x)

# If the supporting SNPs are trans-eqtls reported in eQTLGen

Using top 10 genes with largest abs(gene loadings) and the supporting SNPs, we have 7960 gene ~ SNP pairs. 17 of them were reported by eQTLGen

```{r}

load("/project/xinhe/xsun/pathway_factor/analysis/3.results_analysis/results/trans_pairs.rdata")
DT::datatable(pairs_all[pairs_all$trans,])
```



