---
title: "ACAT resultst"
author: "XSun"
date: "2025-09-17"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---


```{r }
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
suppressMessages(library(tidyr))
suppressMessages(library(forcats))

celltypes <- c("B_cell","CD14_positive_monocyte","CD15_positive_leukocyte","platelet","T_cell","thymocyte")
type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")

qqplot <- function(pvalues,title=NULL) {
  
  pval <- pvalues[complete.cases(pvalues)]
  title <- title
  plotdata <- data.frame(observed = -log10(sort(as.numeric(pval))),
                         expected = -log10(ppoints(length(pval))))
  
  qq <- ggplot(plotdata) + theme_bw(base_line_size =0.3) +
    geom_point(aes(expected, observed), shape = 1, size = 1.5,color = "steelblue") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.6, color = "red") +
    ggtitle(title) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = expression(paste("Expected -log"[10],"(p-value)")),
         y = expression(paste("Observed -log"[10],"(p-value)"))) +
    
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12, color = "black") )
  
  return(qq)
  
}

histogram <- function(pvalues,title=NULL,xlab =NULL) {
  
  p <- as.data.frame(pvalues)
  colnames(p) <- "p"
  
  plot <-  ggplot(p, aes(x=p)) + geom_histogram(breaks = seq(0, 1, by = 0.05),color="white",fill = "steelblue")+ 
    theme_bw(base_line_size =0.3) +
    ggtitle(title) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = xlab,
         y = "Count") +
    
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12, color = "black"),
          text= element_text(family="Times")) 
  
  return(plot)
  
}

qqplot_multi <- function(pvalues_list, legend_names = NULL, colors = NULL, title = NULL) {
  
  # Check if legend names are provided, else use default names
  if (is.null(legend_names) || length(legend_names) != length(pvalues_list)) {
    legend_names <- paste("Set", seq_along(pvalues_list))
  }
  
  # Check if colors are provided, else use rainbow colors
  if (is.null(colors) || length(colors) != length(pvalues_list)) {
    colors <- rainbow(length(pvalues_list))
  }
  
  # Initialize an empty list for storing data frames
  plotdata_list <- list()
  
  # Loop through each vector of p-values
  for (i in seq_along(pvalues_list)) {
    pval <- pvalues_list[[i]][complete.cases(pvalues_list[[i]])]
    data <- data.frame(
      observed = -log10(sort(as.numeric(pval))),
      expected = -log10(ppoints(length(pval))),
      set = legend_names[i]  # Use provided legend name
    )
    plotdata_list[[i]] <- data
  }
  
  # Combine all data frames into one
  plotdata <- do.call(rbind, plotdata_list)
  
  # Plotting
  qq <- ggplot(plotdata, aes(x = expected, y = observed, color = set)) + 
    theme_bw(base_line_size = 0.3) +
    geom_point(shape = 1, size = 1.5) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.6, color = "black") +
    ggtitle(title) + 
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = expression(paste("Expected -log"[10], "(p-value)")),
         y = expression(paste("Observed -log"[10], "(p-value)")),
         color = "Groups") +
    scale_color_manual(values = colors) +  # Use provided or default colors
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12, color = "black"))
  
  return(qq)
}

```

# Original ACAT p-values

## MHC included 

```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p1 <- list()
p2 <- list()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_real_acat.rdata")
  load(file_acat)
  
  p <- unlist(acat_celltype)
  p1[[i]] <- histogram(pvalues = p,title = type[i],xlab = "ACAT p-values from real data")
  p2[[i]] <- qqplot(pvalues = p,title = type[i])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)


```

## MHC removed 

```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p1 <- list()
p2 <- list()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_real_acat_mhcexcluded.rdata")
  load(file_acat)
  
  p <- unlist(acat_celltype)
  p1[[i]] <- histogram(pvalues = p,title = type[i],xlab = "ACAT p-values from real data")
  p2[[i]] <- qqplot(pvalues = p,title = type[i])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```


# Corrected ACAT p-values (pair null)

## MHC included

We sampled the SNPs 100 times, matching the number of SNP in LD & MAF using vSampler. Then we used the random SNPs the do the factor ~ SNP association tests and compute the acat p-values for each pair. We used these random acat p-values to correct the origin ACAT p-values. 

If we corrected the ACAT p-values within each pair, we have: 

(the lowest p-value = 0+1/100+1)

```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p1 <- list()
p2 <- list()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_corrected_acat_pair.rdata")
  load(file_acat)
  
  p <- unlist(acat_p_corrected)
  p1[[i]] <- histogram(pvalues = p,title = type[i],xlab = "Corrected ACAT p-values")
  p2[[i]] <- qqplot(pvalues = p,title = type[i])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```

## MHC removed 

```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p1 <- list()
p2 <- list()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_corrected_acat_pair_mhcexcluded.rdata")
  load(file_acat)
  
  p <- unlist(acat_p_corrected)
  p1[[i]] <- histogram(pvalues = p,title = type[i],xlab = "Corrected ACAT p-values")
  p2[[i]] <- qqplot(pvalues = p,title = type[i])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```



# Corrected ACAT p-values (global null)

## MHC included

If we corrected the ACAT p-values within each cell type (pool all ACAT p-values in each cell type), we have: 

(the lowest p-value = 0+1/(total pair * 100)+1)


```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p1 <- list()
p2 <- list()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_corrected_acat_pool_celltype.rdata")
  load(file_acat)
  
  p <- unlist(acat_p_corrected)
  p1[[i]] <- histogram(pvalues = p,title = type[i],xlab = "Corrected ACAT p-values")
  p2[[i]] <- qqplot(pvalues = p,title = type[i])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```


## MHC removed 

```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p1 <- list()
p2 <- list()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_corrected_acat_pool_celltype_mhcexcluded.rdata")
  load(file_acat)
  
  p <- unlist(acat_p_corrected)
  p1[[i]] <- histogram(pvalues = p,title = type[i],xlab = "Corrected ACAT p-values")
  p2[[i]] <- qqplot(pvalues = p,title = type[i])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```








# Corrected ACAT p-values (trait null)

## MHC included

If we corrected the ACAT p-values within each cell type (pool all ACAT p-values in each cell type), we have: 

(the lowest p-value = 0+1/(total pair * 100)+1)


```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p1 <- list()
p2 <- list()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_corrected_trait_acat.rdata")
  load(file_acat)
  
  p <- unlist(acat_p_corrected)
  p1[[i]] <- histogram(pvalues = p,title = type[i],xlab = "Corrected ACAT p-values")
  p2[[i]] <- qqplot(pvalues = p,title = type[i])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```


## MHC removed 

```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p1 <- list()
p2 <- list()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_corrected_trait_acat_mhcexcluded.rdata")
  load(file_acat)
  
  p <- unlist(acat_p_corrected)
  p1[[i]] <- histogram(pvalues = p,title = type[i],xlab = "Corrected ACAT p-values")
  p2[[i]] <- qqplot(pvalues = p,title = type[i])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```


# All celltypes pooled

## MHC included

```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p_real <- c()
p_random <- c()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_real_acat.rdata")
  load(file_acat)
  p <- unlist(acat_celltype)
  p_real <- c(p_real,p)
  
  file_acat <- paste0(folder_acat,celltypes[i],"_random_acat.rdata")
  load(file_acat)
  p <- unlist(acat_all)
  p_random <- c(p_random,p)

}

set.seed(1)
p_random_sample <- sample(p_random,size = length(p_random)/100,replace = F)

p_list <- list(p_real,p_random_sample)

qqplot_multi(pvalues_list = p_list,legend_names = c("p_real","p_random"))

```



## MHC removed

```{r}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/"

p_real <- c()
p_random <- c()
for (i in 1:length(celltypes)) {
  
  file_acat <- paste0(folder_acat,celltypes[i],"_real_acat_mhcexcluded.rdata")
  load(file_acat)
  p <- unlist(acat_celltype)
  p_real <- c(p_real,p)
  
  file_acat <- paste0(folder_acat,celltypes[i],"_random_acat_mhcexcluded.rdata")
  load(file_acat)
  p <- unlist(acat_all)
  p_random <- c(p_random,p)

}

set.seed(1)
p_random_sample <- sample(p_random,size = length(p_random)/100,replace = F)

p_list <- list(p_real,p_random_sample)

qqplot_multi(pvalues_list = p_list,legend_names = c("p_real","p_random"))

```




<!-- ## corrected  -->

<!-- ```{r} -->

<!-- load("/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/corrected_acat_poolall.rdata") -->
<!-- histogram(pvalues = acat_p_correct,xlab = "Corrected ACAT p-values") -->
<!-- qqplot(acat_p_correct) -->

<!-- ``` -->


# Histogram for Storey's pi1 computed from ACAT p-values

## Origin p-values -- MHC included

```{r} 

load("/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/story_real_acat.rdata")

values <- pi1_all
#labels <- c('CD19+', 'CD14+', 'CD15+', 'PLT', 'CD4+', 'CD8+')
#labels <- names(pi1_all)
labels <- c("B cell \n(CD19+)","Monocyte \n(CD14+)","Leukocyte \n(CD15+)","Platelet","T cell \n(CD4+)","T cell \n(CD8+)")
data <- data.frame(labels, values)

# Create the bar plot
p <- ggplot(data, aes(x=labels, y=values, fill=labels)) +
  geom_bar(stat="identity", color="steelblue", fill="steelblue") +
  labs(x="Cell Types", y="Storey's Pi1 for origin acat p-values") +
  theme_minimal()  +
  
  theme(axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1,vjust = 1.1),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12, color = "black"),
        text= element_text(family="Times")) 

p
```




## Origin p-values -- MHC removed

```{r} 

load("/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/story_real_acat_mhcexcluded.rdata")

values <- pi1_all
#labels <- c('CD19+', 'CD14+', 'CD15+', 'PLT', 'CD4+', 'CD8+')
#labels <- names(pi1_all)
labels <- c("B cell \n(CD19+)","Monocyte \n(CD14+)","Leukocyte \n(CD15+)","Platelet","T cell \n(CD4+)","T cell \n(CD8+)")
data <- data.frame(labels, values)

# Create the bar plot
p <- ggplot(data, aes(x=labels, y=values, fill=labels)) +
  geom_bar(stat="identity", color="steelblue", fill="steelblue") +
  labs(x="Cell Types", y="Storey's Pi1 for origin acat p-values") +
  theme_minimal()  +
  
  theme(axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1,vjust = 1.1),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12, color = "black"),
        text= element_text(family="Times")) 

p
```


## Corrected p-values -- Global null -- MHC included

If we corrected the ACAT p-values within each pair


```{r} 

load("/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/story_corrected_acat.rdata")

values <- pi1_all
#labels <- c('CD19+', 'CD14+', 'CD15+', 'PLT', 'CD4+', 'CD8+')
#labels <- names(pi1_all)
labels <- c("B cell \n(CD19+)","Monocyte \n(CD14+)","Leukocyte \n(CD15+)","Platelet","T cell \n(CD4+)","T cell \n(CD8+)")
data <- data.frame(labels, values)

# Create the bar plot
p <- ggplot(data, aes(x=labels, y=values, fill=labels)) +
  geom_bar(stat="identity", color="steelblue", fill="steelblue") +
  labs(x="Cell Types", y="Storey's Pi1 for corrected acat p-values") +
  theme_minimal()  +
  
  theme(axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1,vjust = 1.1),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12, color = "black"),
        text= element_text(family="Times")) 

p
```

If we corrected the ACAT p-values within each cell type (pool all ACAT p-values in each cell type)


```{r} 

load("/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/story_corrected_acat_pool_celltype.rdata")

values <- pi1_all
#labels <- c('CD19+', 'CD14+', 'CD15+', 'PLT', 'CD4+', 'CD8+')
#labels <- names(pi1_all)
labels <- c("B cell \n(CD19+)","Monocyte \n(CD14+)","Leukocyte \n(CD15+)","Platelet","T cell \n(CD4+)","T cell \n(CD8+)")
data <- data.frame(labels, values)

# Create the bar plot
p <- ggplot(data, aes(x=labels, y=values, fill=labels)) +
  geom_bar(stat="identity", color="steelblue", fill="steelblue") +
  labs(x="Cell Types", y="Storey's Pi1 for corrected acat p-values") +
  theme_minimal()  +
  
  theme(axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1,vjust = 1.1),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12, color = "black"),
        text= element_text(family="Times")) 

p
```


## Corrected p-values -- Global null - MHC removed 

If we corrected the ACAT p-values within each pair

```{r} 

load("/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/story_corrected_acat_mhcexcluded.rdata")

values <- pi1_all
#labels <- c('CD19+', 'CD14+', 'CD15+', 'PLT', 'CD4+', 'CD8+')
#labels <- names(pi1_all)
labels <- c("B cell \n(CD19+)","Monocyte \n(CD14+)","Leukocyte \n(CD15+)","Platelet","T cell \n(CD4+)","T cell \n(CD8+)")
data <- data.frame(labels, values)

# Create the bar plot
p <- ggplot(data, aes(x=labels, y=values, fill=labels)) +
  geom_bar(stat="identity", color="steelblue", fill="steelblue") +
  labs(x="Cell Types", y="Storey's Pi1 for corrected acat p-values") +
  theme_minimal()  +
  
  theme(axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1,vjust = 1.1),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12, color = "black"),
        text= element_text(family="Times")) 

p
```

If we corrected the ACAT p-values within each cell type (pool all ACAT p-values in each cell type)


```{r} 

load("/project/xinhe/xsun/pathway_factor/analysis/2.ACAT/results/story_corrected_acat_pool_celltype_mhcexcluded.rdata")

values <- pi1_all
#labels <- c('CD19+', 'CD14+', 'CD15+', 'PLT', 'CD4+', 'CD8+')
#labels <- names(pi1_all)
labels <- c("B cell \n(CD19+)","Monocyte \n(CD14+)","Leukocyte \n(CD15+)","Platelet","T cell \n(CD4+)","T cell \n(CD8+)")
data <- data.frame(labels, values)

# Create the bar plot
p <- ggplot(data, aes(x=labels, y=values, fill=labels)) +
  geom_bar(stat="identity", color="steelblue", fill="steelblue") +
  labs(x="Cell Types", y="Storey's Pi1 for corrected acat p-values") +
  theme_minimal()  +
  
  theme(axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1,vjust = 1.1),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12, color = "black"),
        text= element_text(family="Times")) 

p
```