---
title: "Figures"
author: "XSun"
date: "2025-11-05"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

```{r, warning=F, message=F}

library(gridExtra)
library(ggplot2)
library(dplyr)

celltypes <- c("B_cell","CD14_positive_monocyte","CD15_positive_leukocyte","platelet","T_cell","thymocyte")
types <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")
names(celltypes) <- types
names(types) <- celltypes

source("/project/xinhe/xsun/pathway_factor/data_v2/traits_finalselection_SLEremoved.R")


qqplot <- function(pvalues,title=NULL) {
  
  pval <- pvalues[complete.cases(pvalues)]
  title <- title
  plotdata <- data.frame(observed = -log10(sort(as.numeric(pval))),
                         expected = -log10(ppoints(length(pval))))
  
  qq <- ggplot(plotdata) + theme_bw(base_line_size =0.3) +
    geom_point(aes(expected, observed), shape = 1, size = 1.5,color = "steelblue") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.6, color = "red") +
    ggtitle(title) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = expression(paste("Expected -log"[10],"(p-value)")),
         y = expression(paste("Observed -log"[10],"(p-value)"))) +
    
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12, color = "black"),
          text= element_text(family="Times"))
  
  return(qq)
  
}

```


# SI figures

The main change for figure S1-S3 is: more factors are included since we decreased the gene num cutoff from 40 to 20.

## S1 hsq and SE

```{r fig.height=5, fig.width=7.7, warning=F, message=F}

p1 <- list()
p2 <- list()
for (celltype in celltypes){
  
  file_hsq <- paste0("/project/xinhe/xsun/pathway_factor/analysis/7.analysis_for_factors/results/heritability/hsq_summary_", celltype,".txt")
  
  H <- read.table(file_hsq, header = T)
  sum_sorted <- H[order(as.numeric(H$h2)),]
  sum_sorted$order <- seq(1,nrow(H),by=1)
  
  p1[[celltype]] <- ggplot(sum_sorted, aes(x=order, y=h2)) + theme_bw(base_line_size =0.3) +
    geom_errorbar(aes(ymin=h2-SE, ymax=h2+SE), width=0.1, color = "grey",alpha=0.6) +
    geom_point(size = 0.8, color = "steelblue") + 
    ggtitle(types[celltype]) + theme(plot.title = element_text(hjust = 0.5)) + 
    xlab("PCs sorted by heritability") + ylab("Heritability ± SE") + 
    ylim(-2,3)  + 
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12, color = "black"),
          text= element_text(family="Times")) 
  
  plotdata <- data.frame(observed = -log10(sort(H$Pval)),
                         expected = -log10(ppoints(nrow(H))))
  
  
  p2[[celltype]] <- ggplot(plotdata) + theme_bw(base_line_size =0.3) +
    geom_point(aes(expected, observed), shape = 1, size = 1.5, color = "steelblue") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.6, color = "red") +
    ggtitle(types[celltype]) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = expression(paste("Expected -log"[10],"(p-value)")),
         y = expression(paste("Observed -log"[10],"(p-value)"))) +
    
    theme(axis.title.x = element_text(size = 10),
          axis.text.x = element_text(size = 8, color = "black"),
          axis.title.y = element_text(size = 10),
          axis.text.y = element_text(size = 8, color = "black"),
          text= element_text(family="Times")) 
  
  
  
  
}


all_interval <- grid.arrange(p1[[1]],p1[[2]],p1[[3]],p1[[4]],p1[[5]],p1[[6]], nrow = 2)

```


## S2 hsq qqplot for p-values

```{r fig.height=5, fig.width=7.7, warning=F, message=F}


all_qq <- grid.arrange(p2[[1]],p2[[2]],p2[[3]],p2[[4]],p2[[5]],p2[[6]], nrow = 2)


```


<!-- ## S3 Number of independent signals for factor gwas (factors filtered by hsq >0.9 & hsq pvalue < 0.05) -->


<!-- ```{r fig.height=5, fig.width=7.7, warning=F, message=F} -->


<!-- p1 <- list() -->
<!-- for (i in 1:length(types)){  -->

<!--   file <- paste0("/project/xinhe/xsun/pathway_factor/analysis/7.analysis_for_factors/results/num_gwas_signals_", celltype, ".RDS") -->
<!--   num_signal_all <- readRDS(file) -->

<!--   num_signal_all <- as.data.frame(num_signal_all) -->
<!--   colnames(num_signal_all) <- "num" -->
<!--   #p1[[i]] <-  ggplot(num_signal_all, aes(x=num)) + geom_histogram(breaks = seq(0, max(num_signal_all), by = 1),color="white",fill = "steelblue")+ -->
<!--   p1[[i]] <-  ggplot(num_signal_all, aes(x=num)) + geom_histogram(bins = 30,color="white",fill = "steelblue")+ -->
<!--     theme_bw(base_line_size =0.3) + -->
<!--     ggtitle(types[i]) + theme(plot.title = element_text(hjust = 0.5)) + -->
<!--     labs(x = "# of independent signals", -->
<!--          y = "Count") + -->

<!--     theme(axis.title.x = element_text(size = 14), -->
<!--           axis.text.x = element_text(size = 12, color = "black"), -->
<!--           axis.title.y = element_text(size = 14), -->
<!--           axis.text.y = element_text(size = 12, color = "black"), -->
<!--           text= element_text(family="Times")) + -->

<!--     annotate(geom = "text", x = Inf, y = Inf, label = paste0("The total number of tested factor = ",nrow(num_signal_all)), -->
<!--              hjust = 1, vjust = 1,  size = 2,  family = "Times",   color = "black") -->


<!-- } -->

<!-- all1 <- grid.arrange(p1[[1]],p1[[2]],p1[[3]],p1[[4]],p1[[5]],p1[[6]], nrow = 2) -->


<!-- ``` -->




## S4 2 examples for ACAT

same examples with initial submission. but the snps are a bit different

```{r fig.height=3, fig.width=5.5, warning=F, message=F}


folder_harmo <- "/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/assoc_harmonize/"


celltype <- "B_cell"
trait <- "lymphc-ebi-a-GCST90002388"
factor <- "04014_PC3"

file_harmo <- paste0(folder_harmo,celltype,"-",trait,"_harmo.RDS")
list_harmo <- readRDS(file_harmo)

df_harmo <- list_harmo[[factor]]

p1 <- qqplot(df_harmo$pval.exposure, title = "3rd PC of Ras signaling pathway \n- lymphocyte count")


celltype <- "B_cell"
trait <- "IBD-ebi-a-GCST003043"
factor <- "04612_PC5"

file_harmo <- paste0(folder_harmo,celltype,"-",trait,"_harmo.RDS")
list_harmo <- readRDS(file_harmo)

df_harmo <- list_harmo[[factor]]

p2 <- qqplot(df_harmo$pval.exposure, title = "5th PC of Antigen processing and \npresentation pathway - IBD")


all <- grid.arrange(p1,p2, nrow = 1)


```

## S5 Number of supporting SNPs per pair

difference: the current cutoff is >=5SNPs and >=20genes

### EUR


```{r fig.height=5, fig.width=7.7, warning=F, message=F}

p <- list()
for(celltype in celltypes){
  
  df <- readRDS(paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/ECT_summary_final/", celltype,"_ECT_summary_EUR.RDS"))
  df <- df[df$trait %in% traits_EUR,]
  
  summary <- df %>%
    count(num_supp_SNP, name = "Freq")
  
  # Plot
  p[[celltype]] <- ggplot(summary, aes(x = num_supp_SNP, y = Freq)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    theme_bw(base_line_size = 0.3) +
    ggtitle(types[celltype]) +
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = "# of supporting SNPs in each pair", y = "Count") +
    geom_text(
      aes(label = ifelse(num_supp_SNP %% 5 == 0, as.character(Freq), "")),
      colour = "black", size = 2, vjust = -0.1, family = "Times"
    ) +
    annotate(
      geom = "text",
      x = Inf, y = Inf,
      label = paste0("# of candidate pairs = ", sum(summary$Freq[summary$num_supp_SNP >= 5])),
      hjust = 1.1, vjust = 1.5, size = 3, family = "Times", color = "black"
    ) +
    theme(
      axis.title.x = element_text(size = 10),
      axis.text.x  = element_text(size = 8, color = "black"),
      axis.title.y = element_text(size = 10),
      axis.text.y  = element_text(size = 8, color = "black"),
      text = element_text(family = "Times")
    ) +
    scale_x_continuous(breaks = seq(0, max(summary$num_supp_SNP), 5))
  
}


all <- grid.arrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]],p[[6]], nrow = 2)



```



### EAS


```{r fig.height=5, fig.width=7.7, warning=F, message=F}

p <- list()
for(celltype in celltypes){
  
  df <- readRDS(paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/ECT_summary_final/", celltype,"_ECT_summary_EAS.RDS"))
  df <- df[df$trait %in% traits_EAS,]
  
  summary <- df %>%
    count(num_supp_SNP, name = "Freq")
  
  # Plot
  p[[celltype]] <- ggplot(summary, aes(x = num_supp_SNP, y = Freq)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    theme_bw(base_line_size = 0.3) +
    ggtitle(types[celltype]) +
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = "# of supporting SNPs in each pair", y = "Count") +
    geom_text(
      aes(label = ifelse(num_supp_SNP %% 5 == 0, as.character(Freq), "")),
      colour = "black", size = 2, vjust = -0.1, family = "Times"
    ) +
    annotate(
      geom = "text",
      x = Inf, y = Inf,
      label = paste0("# of candidate pairs = ", sum(summary$Freq[summary$num_supp_SNP >= 5])),
      hjust = 1.1, vjust = 1.5, size = 3, family = "Times", color = "black"
    ) +
    theme(
      axis.title.x = element_text(size = 10),
      axis.text.x  = element_text(size = 8, color = "black"),
      axis.title.y = element_text(size = 10),
      axis.text.y  = element_text(size = 8, color = "black"),
      text = element_text(family = "Times")
    ) +
    scale_x_continuous(breaks = seq(0, max(summary$num_supp_SNP), 5))
  
}


all <- grid.arrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]],p[[6]], nrow = 2)



```


## S6 ECT p-values

### EUR

```{r fig.height=5, fig.width=7.7, warning=F, message=F}

p <- list()
for(celltype in celltypes){
  
  df <- readRDS(paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/ECT_summary_final/", celltype,"_ECT_summary_EUR.RDS"))
  df <- df[df$trait %in% traits_EUR,]
  
  df <- df[complete.cases(df$p_ECT),]
  
  summary_p <- df %>%
    mutate(bin = cut(p_ECT, breaks = seq(0, 1, by = 0.05), include.lowest = TRUE)) %>%
    count(bin, name = "Freq")
  
  # Correctly handle both [ and ( at the start of interval labels
  summary_p <- summary_p %>%
    mutate(
      bin_low  = as.numeric(sub("[^0-9\\.]*([0-9\\.]+),.*", "\\1", bin)),
      bin_high = as.numeric(sub(".*,([0-9\\.]+)\\]", "\\1", bin)),
      bin_mid  = (bin_low + bin_high) / 2
    )
  
  p[[celltype]] <- ggplot(summary_p, aes(x = bin_mid, y = Freq)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    theme_bw(base_line_size = 0.3) +
    ggtitle(types[celltype]) +
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = "ECT p-value", y = "Count") +
    theme(
      axis.title.x = element_text(size = 10),
      axis.text.x  = element_text(size = 8, color = "black"),
      axis.title.y = element_text(size = 10),
      axis.text.y  = element_text(size = 8, color = "black"),
      text = element_text(family = "Times")
    ) +
    scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
  
}

all <- grid.arrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]],p[[6]], nrow = 2)

p <- list()
for(celltype in celltypes){
  
  df <- readRDS(paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/ECT_summary_final/", celltype,"_ECT_summary_EUR.RDS"))
  df <- df[df$trait %in% traits_EUR,]
  
  df <- df[complete.cases(df$p_ECT),]
  
  p[[celltype]] <- qqplot(pvalues = df$p_ECT, title = types[celltype])
  
}

all <- grid.arrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]],p[[6]], nrow = 2)


```





### EAS

```{r fig.height=5, fig.width=7.7, warning=F, message=F}

p <- list()
for(celltype in celltypes){
  
  df <- readRDS(paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/ECT_summary_final/", celltype,"_ECT_summary_EAS.RDS"))
  df <- df[df$trait %in% traits_EAS,]
  
  df <- df[complete.cases(df$p_ECT),]
  
  summary_p <- df %>%
    mutate(bin = cut(p_ECT, breaks = seq(0, 1, by = 0.05), include.lowest = TRUE)) %>%
    count(bin, name = "Freq")
  
  # Correctly handle both [ and ( at the start of interval labels
  summary_p <- summary_p %>%
    mutate(
      bin_low  = as.numeric(sub("[^0-9\\.]*([0-9\\.]+),.*", "\\1", bin)),
      bin_high = as.numeric(sub(".*,([0-9\\.]+)\\]", "\\1", bin)),
      bin_mid  = (bin_low + bin_high) / 2
    )
  
  p[[celltype]] <- ggplot(summary_p, aes(x = bin_mid, y = Freq)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    theme_bw(base_line_size = 0.3) +
    ggtitle(types[celltype]) +
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = "ECT p-value", y = "Count") +
    theme(
      axis.title.x = element_text(size = 10),
      axis.text.x  = element_text(size = 8, color = "black"),
      axis.title.y = element_text(size = 10),
      axis.text.y  = element_text(size = 8, color = "black"),
      text = element_text(family = "Times")
    ) +
    scale_x_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1))
  
}

all <- grid.arrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]],p[[6]], nrow = 2)

p <- list()
for(celltype in celltypes){
  
  df <- readRDS(paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/ECT_summary_final/", celltype,"_ECT_summary_EAS.RDS"))
  df <- df[df$trait %in% traits_EAS,]
  
  df <- df[complete.cases(df$p_ECT),]
  
  p[[celltype]] <- qqplot(pvalues = df$p_ECT, title = types[celltype])
  
}

all <- grid.arrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]],p[[6]], nrow = 2)


```



## S7 ECT storey pi1

### EUR


```{r fig.height=3, fig.width=5, warning=F, message=F}

load("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/results/ECT_story_EUR.rdata")

values <- pi1_all
labels <- c("B cell \n(CD19+)","Monocyte \n(CD14+)","Leukocyte \n(CD15+)","Platelet","T cell \n(CD4+)","T cell \n(CD8+)")
data <- data.frame(labels, values)

# Create the bar plot
p <- ggplot(data, aes(x=labels, y=values, fill=labels)) +
  geom_bar(stat="identity", color="steelblue", fill="steelblue") +
  labs(x="Cell Types", y="Storey's Pi1") +
  theme_minimal()  +
  
  theme(axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1,vjust = 1.1),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12, color = "black"),
        text= element_text(family="Times")) 

p

```

### EAS


```{r fig.height=3, fig.width=5, warning=F, message=F}

load("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/results/ECT_story_EAS.rdata")

values <- pi1_all
labels <- c("B cell \n(CD19+)","Monocyte \n(CD14+)","Leukocyte \n(CD15+)","Platelet","T cell \n(CD4+)","T cell \n(CD8+)")
data <- data.frame(labels, values)

# Create the bar plot
p <- ggplot(data, aes(x=labels, y=values, fill=labels)) +
  geom_bar(stat="identity", color="steelblue", fill="steelblue") +
  labs(x="Cell Types", y="Storey's Pi1") +
  theme_minimal()  +
  
  theme(axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12, color = "black", angle = 45, hjust = 1,vjust = 1.1),
        axis.title.y = element_text(size = 14),
        axis.text.y = element_text(size = 12, color = "black"),
        text= element_text(family="Times")) 

p

```