---
title: "ECT results"
author: "XSun"
date: "2025-10-10"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true  
---


|           Full Trait Name           |   Abbr.   |             Category             |
|:-----------------------------------:|:---------:|:--------------------------------:|
| Crohn’s disease                     | CD        | Immune / Inflammatory disease    |
| Inflammatory bowel disease          | IBD       | Immune / Inflammatory disease    |
| Rheumatoid arthritis                | RA        | Immune / Inflammatory disease    |
| Systemic lupus erythematosus        | SLE       | Immune / Inflammatory disease    |
| Psoriasis                           | Psoriasis | Immune / Inflammatory disease    |
| Asthma                              | asthma    | Immune / Inflammatory disease    |
| Multiple sclerosis                  | MS        | Immune / Neurological disease    |
| Celiac disease                      | celiac    | Immune / Inflammatory disease    |
| Hashimoto thyroiditis               | HT        | Immune / Endocrine disease       |
| Type 1 diabetes                     | T1D       | Immune / Metabolic disease       |
| Ankylosing spondylitis              | AS        | Immune / Inflammatory disease    |
| Allergy                             | ALLERGY   | Immune / Hypersensitivity        |
| Hemoglobin concentration            | HB        | Hematological trait              |
| Hematocrit                          | HCT       | Hematological trait              |
| Mean corpuscular volume             | MCV       | Hematological trait              |
| Mean corpuscular hemoglobin         | MCH       | Hematological trait              |
| Platelet count                      | PLT       | Hematological trait              |
| Red blood cell count                | RBC       | Hematological trait              |
| White blood cell count              | WBC       | Hematological trait              |
| Low-density lipoprotein cholesterol | LDL       | Metabolic / Cardiovascular trait |



```{r}

library(tidyverse)
library(ggplot2)

celltypes <- c("B_cell","CD14_positive_monocyte","CD15_positive_leukocyte","platelet","T_cell","thymocyte")
type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")
names(type) <- celltypes

source("/project/xinhe/xsun/pathway_factor/data_v2/traits_selected.R")

folder_ect <- "/project/xinhe/xsun/pathway_factor/analysis/1.ECT_v2/ECT_summary/"



```

# 40 genes

## Old GWAS traits

```{r fig.width=8, fig.height=5}

df_all_FDR01 <- c()
df_all_FDR02 <- c()
df_p0001 <- c()
for (celltype in celltypes){
  
  df <- readRDS(paste0(folder_ect,celltype,"_ECT_summary_40genes_selected_old.RDS"))
  
  tmp_FDR01 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.1,na.rm = T),sum(df$FDR_4suppSNP < 0.1,na.rm = T),sum(df$FDR_5suppSNP < 0.1,na.rm = T),sum(df$FDR_6suppSNP < 0.1,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.1,na.rm = T),sum(df$FDR_8suppSNP < 0.1,na.rm = T),sum(df$FDR_9suppSNP < 0.1,na.rm = T),sum(df$FDR_10suppSNP < 0.1,na.rm = T))
  tmp_FDR02 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.2,na.rm = T),sum(df$FDR_4suppSNP < 0.2,na.rm = T),sum(df$FDR_5suppSNP < 0.2,na.rm = T),sum(df$FDR_6suppSNP < 0.2,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.2,na.rm = T),sum(df$FDR_8suppSNP < 0.2,na.rm = T),sum(df$FDR_9suppSNP < 0.2,na.rm = T),sum(df$FDR_10suppSNP < 0.2,na.rm = T))
  tmp_FDR02[2:length(tmp_FDR02)] <- as.numeric(tmp_FDR02[2:length(tmp_FDR02)]) - as.numeric(tmp_FDR01[2:length(tmp_FDR01)])
  
  df_all_FDR01 <- rbind(df_all_FDR01,tmp_FDR01)
  df_all_FDR02 <- rbind(df_all_FDR02,tmp_FDR02)
  
  df_p0001 <- rbind(df_p0001, df[df$p_ECT < 0.001,])
  
}

df_p0001 <- df_p0001[complete.cases(df_p0001$p_ECT),]

rownames(df_all_FDR01) <- NULL
colnames(df_all_FDR01) <- c("celltype","num_pairs_FDR01_3suppSNP","num_pairs_FDR01_4suppSNP","num_pairs_FDR01_5suppSNP",
                            "num_pairs_FDR01_6suppSNP","num_pairs_FDR01_7suppSNP","num_pairs_FDR01_8suppSNP",
                            "num_pairs_FDR01_9suppSNP","num_pairs_FDR01_10suppSNP")
df_all_FDR01 <- as.data.frame(df_all_FDR01, stringsAsFactors = FALSE)
df_all_FDR01 <- df_all_FDR01 %>% mutate(across(-celltype, as.numeric))

df_long <- df_all_FDR01 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR01_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR01_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.1)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )



rownames(df_all_FDR02) <- NULL
colnames(df_all_FDR02) <- c("celltype","num_pairs_FDR02_3suppSNP","num_pairs_FDR02_4suppSNP","num_pairs_FDR02_5suppSNP",
                            "num_pairs_FDR02_6suppSNP","num_pairs_FDR02_7suppSNP","num_pairs_FDR02_8suppSNP",
                            "num_pairs_FDR02_9suppSNP","num_pairs_FDR02_10suppSNP")
df_all_FDR02 <- as.data.frame(df_all_FDR02, stringsAsFactors = FALSE)
df_all_FDR02 <- df_all_FDR02 %>% mutate(across(-celltype, as.numeric))


df_long <- df_all_FDR02 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR02_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR02_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.2)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )


DT::datatable(df_p0001,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','significant pairs '),options = list(pageLength = 10) )



```


## New EUR traits

```{r fig.width=8, fig.height=5}

df_all_FDR01 <- c()
df_all_FDR02 <- c()
df_p0001 <- c()
for (celltype in celltypes){
  
  df <- readRDS(paste0(folder_ect,celltype,"_ECT_summary_40genes_selected_newbatch.RDS"))
  
  tmp_FDR01 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.1,na.rm = T),sum(df$FDR_4suppSNP < 0.1,na.rm = T),sum(df$FDR_5suppSNP < 0.1,na.rm = T),sum(df$FDR_6suppSNP < 0.1,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.1,na.rm = T),sum(df$FDR_8suppSNP < 0.1,na.rm = T),sum(df$FDR_9suppSNP < 0.1,na.rm = T),sum(df$FDR_10suppSNP < 0.1,na.rm = T))
  tmp_FDR02 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.2,na.rm = T),sum(df$FDR_4suppSNP < 0.2,na.rm = T),sum(df$FDR_5suppSNP < 0.2,na.rm = T),sum(df$FDR_6suppSNP < 0.2,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.2,na.rm = T),sum(df$FDR_8suppSNP < 0.2,na.rm = T),sum(df$FDR_9suppSNP < 0.2,na.rm = T),sum(df$FDR_10suppSNP < 0.2,na.rm = T))
  tmp_FDR02[2:length(tmp_FDR02)] <- as.numeric(tmp_FDR02[2:length(tmp_FDR02)]) - as.numeric(tmp_FDR01[2:length(tmp_FDR01)])
  
  df_all_FDR01 <- rbind(df_all_FDR01,tmp_FDR01)
  df_all_FDR02 <- rbind(df_all_FDR02,tmp_FDR02)
  
  df_p0001 <- rbind(df_p0001, df[df$p_ECT < 0.001,])
  
}

df_p0001 <- df_p0001[complete.cases(df_p0001$p_ECT),]

rownames(df_all_FDR01) <- NULL
colnames(df_all_FDR01) <- c("celltype","num_pairs_FDR01_3suppSNP","num_pairs_FDR01_4suppSNP","num_pairs_FDR01_5suppSNP",
                            "num_pairs_FDR01_6suppSNP","num_pairs_FDR01_7suppSNP","num_pairs_FDR01_8suppSNP",
                            "num_pairs_FDR01_9suppSNP","num_pairs_FDR01_10suppSNP")
df_all_FDR01 <- as.data.frame(df_all_FDR01, stringsAsFactors = FALSE)
df_all_FDR01 <- df_all_FDR01 %>% mutate(across(-celltype, as.numeric))

df_long <- df_all_FDR01 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR01_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR01_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.1)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )



rownames(df_all_FDR02) <- NULL
colnames(df_all_FDR02) <- c("celltype","num_pairs_FDR02_3suppSNP","num_pairs_FDR02_4suppSNP","num_pairs_FDR02_5suppSNP",
                            "num_pairs_FDR02_6suppSNP","num_pairs_FDR02_7suppSNP","num_pairs_FDR02_8suppSNP",
                            "num_pairs_FDR02_9suppSNP","num_pairs_FDR02_10suppSNP")
df_all_FDR02 <- as.data.frame(df_all_FDR02, stringsAsFactors = FALSE)
df_all_FDR02 <- df_all_FDR02 %>% mutate(across(-celltype, as.numeric))


df_long <- df_all_FDR02 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR02_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR02_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.2)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )


DT::datatable(df_p0001,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','significant pairs '),options = list(pageLength = 10) )



```


## New EAS traits

```{r fig.width=8, fig.height=5}

df_all_FDR01 <- c()
df_all_FDR02 <- c()
df_p0001 <- c()
for (celltype in celltypes){
  
  df <- readRDS(paste0(folder_ect,celltype,"_ECT_summary_40genes_EAS.RDS"))
  
  tmp_FDR01 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.1,na.rm = T),sum(df$FDR_4suppSNP < 0.1,na.rm = T),sum(df$FDR_5suppSNP < 0.1,na.rm = T),sum(df$FDR_6suppSNP < 0.1,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.1,na.rm = T),sum(df$FDR_8suppSNP < 0.1,na.rm = T),sum(df$FDR_9suppSNP < 0.1,na.rm = T),sum(df$FDR_10suppSNP < 0.1,na.rm = T))
  tmp_FDR02 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.2,na.rm = T),sum(df$FDR_4suppSNP < 0.2,na.rm = T),sum(df$FDR_5suppSNP < 0.2,na.rm = T),sum(df$FDR_6suppSNP < 0.2,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.2,na.rm = T),sum(df$FDR_8suppSNP < 0.2,na.rm = T),sum(df$FDR_9suppSNP < 0.2,na.rm = T),sum(df$FDR_10suppSNP < 0.2,na.rm = T))
  tmp_FDR02[2:length(tmp_FDR02)] <- as.numeric(tmp_FDR02[2:length(tmp_FDR02)]) - as.numeric(tmp_FDR01[2:length(tmp_FDR01)])
  
  df_all_FDR01 <- rbind(df_all_FDR01,tmp_FDR01)
  df_all_FDR02 <- rbind(df_all_FDR02,tmp_FDR02)
  
  df_p0001 <- rbind(df_p0001, df[df$p_ECT < 0.001,])
  
}

df_p0001 <- df_p0001[complete.cases(df_p0001$p_ECT),]

rownames(df_all_FDR01) <- NULL
colnames(df_all_FDR01) <- c("celltype","num_pairs_FDR01_3suppSNP","num_pairs_FDR01_4suppSNP","num_pairs_FDR01_5suppSNP",
                            "num_pairs_FDR01_6suppSNP","num_pairs_FDR01_7suppSNP","num_pairs_FDR01_8suppSNP",
                            "num_pairs_FDR01_9suppSNP","num_pairs_FDR01_10suppSNP")
df_all_FDR01 <- as.data.frame(df_all_FDR01, stringsAsFactors = FALSE)
df_all_FDR01 <- df_all_FDR01 %>% mutate(across(-celltype, as.numeric))

df_long <- df_all_FDR01 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR01_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR01_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.1)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )



rownames(df_all_FDR02) <- NULL
colnames(df_all_FDR02) <- c("celltype","num_pairs_FDR02_3suppSNP","num_pairs_FDR02_4suppSNP","num_pairs_FDR02_5suppSNP",
                            "num_pairs_FDR02_6suppSNP","num_pairs_FDR02_7suppSNP","num_pairs_FDR02_8suppSNP",
                            "num_pairs_FDR02_9suppSNP","num_pairs_FDR02_10suppSNP")
df_all_FDR02 <- as.data.frame(df_all_FDR02, stringsAsFactors = FALSE)
df_all_FDR02 <- df_all_FDR02 %>% mutate(across(-celltype, as.numeric))


df_long <- df_all_FDR02 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR02_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR02_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.2)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )


DT::datatable(df_p0001,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','significant pairs '),options = list(pageLength = 10) )



```



# 20 genes
## Old GWAS traits

```{r fig.width=8, fig.height=5}

df_all_FDR01 <- c()
df_all_FDR02 <- c()
df_p0001 <- c()
for (celltype in celltypes){
  
  df <- readRDS(paste0(folder_ect,celltype,"_ECT_summary_20genes_selected_old.RDS"))
  
  tmp_FDR01 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.1,na.rm = T),sum(df$FDR_4suppSNP < 0.1,na.rm = T),sum(df$FDR_5suppSNP < 0.1,na.rm = T),sum(df$FDR_6suppSNP < 0.1,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.1,na.rm = T),sum(df$FDR_8suppSNP < 0.1,na.rm = T),sum(df$FDR_9suppSNP < 0.1,na.rm = T),sum(df$FDR_10suppSNP < 0.1,na.rm = T))
  tmp_FDR02 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.2,na.rm = T),sum(df$FDR_4suppSNP < 0.2,na.rm = T),sum(df$FDR_5suppSNP < 0.2,na.rm = T),sum(df$FDR_6suppSNP < 0.2,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.2,na.rm = T),sum(df$FDR_8suppSNP < 0.2,na.rm = T),sum(df$FDR_9suppSNP < 0.2,na.rm = T),sum(df$FDR_10suppSNP < 0.2,na.rm = T))
  tmp_FDR02[2:length(tmp_FDR02)] <- as.numeric(tmp_FDR02[2:length(tmp_FDR02)]) - as.numeric(tmp_FDR01[2:length(tmp_FDR01)])
  
  df_all_FDR01 <- rbind(df_all_FDR01,tmp_FDR01)
  df_all_FDR02 <- rbind(df_all_FDR02,tmp_FDR02)
  
  df_p0001 <- rbind(df_p0001, df[df$p_ECT < 0.001,])
  
}

df_p0001 <- df_p0001[complete.cases(df_p0001$p_ECT),]

rownames(df_all_FDR01) <- NULL
colnames(df_all_FDR01) <- c("celltype","num_pairs_FDR01_3suppSNP","num_pairs_FDR01_4suppSNP","num_pairs_FDR01_5suppSNP",
                            "num_pairs_FDR01_6suppSNP","num_pairs_FDR01_7suppSNP","num_pairs_FDR01_8suppSNP",
                            "num_pairs_FDR01_9suppSNP","num_pairs_FDR01_10suppSNP")
df_all_FDR01 <- as.data.frame(df_all_FDR01, stringsAsFactors = FALSE)
df_all_FDR01 <- df_all_FDR01 %>% mutate(across(-celltype, as.numeric))

df_long <- df_all_FDR01 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR01_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR01_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.1)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )



rownames(df_all_FDR02) <- NULL
colnames(df_all_FDR02) <- c("celltype","num_pairs_FDR02_3suppSNP","num_pairs_FDR02_4suppSNP","num_pairs_FDR02_5suppSNP",
                            "num_pairs_FDR02_6suppSNP","num_pairs_FDR02_7suppSNP","num_pairs_FDR02_8suppSNP",
                            "num_pairs_FDR02_9suppSNP","num_pairs_FDR02_10suppSNP")
df_all_FDR02 <- as.data.frame(df_all_FDR02, stringsAsFactors = FALSE)
df_all_FDR02 <- df_all_FDR02 %>% mutate(across(-celltype, as.numeric))


df_long <- df_all_FDR02 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR02_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR02_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.2)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )


DT::datatable(df_p0001,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','significant pairs '),options = list(pageLength = 10) )



```


## New EUR traits

```{r fig.width=8, fig.height=5}

df_all_FDR01 <- c()
df_all_FDR02 <- c()
df_p0001 <- c()
for (celltype in celltypes){
  
  df <- readRDS(paste0(folder_ect,celltype,"_ECT_summary_20genes_selected_newbatch.RDS"))
  
  tmp_FDR01 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.1,na.rm = T),sum(df$FDR_4suppSNP < 0.1,na.rm = T),sum(df$FDR_5suppSNP < 0.1,na.rm = T),sum(df$FDR_6suppSNP < 0.1,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.1,na.rm = T),sum(df$FDR_8suppSNP < 0.1,na.rm = T),sum(df$FDR_9suppSNP < 0.1,na.rm = T),sum(df$FDR_10suppSNP < 0.1,na.rm = T))
  tmp_FDR02 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.2,na.rm = T),sum(df$FDR_4suppSNP < 0.2,na.rm = T),sum(df$FDR_5suppSNP < 0.2,na.rm = T),sum(df$FDR_6suppSNP < 0.2,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.2,na.rm = T),sum(df$FDR_8suppSNP < 0.2,na.rm = T),sum(df$FDR_9suppSNP < 0.2,na.rm = T),sum(df$FDR_10suppSNP < 0.2,na.rm = T))
  tmp_FDR02[2:length(tmp_FDR02)] <- as.numeric(tmp_FDR02[2:length(tmp_FDR02)]) - as.numeric(tmp_FDR01[2:length(tmp_FDR01)])
  
  df_all_FDR01 <- rbind(df_all_FDR01,tmp_FDR01)
  df_all_FDR02 <- rbind(df_all_FDR02,tmp_FDR02)
  
  df_p0001 <- rbind(df_p0001, df[df$p_ECT < 0.001,])
  
}

df_p0001 <- df_p0001[complete.cases(df_p0001$p_ECT),]

rownames(df_all_FDR01) <- NULL
colnames(df_all_FDR01) <- c("celltype","num_pairs_FDR01_3suppSNP","num_pairs_FDR01_4suppSNP","num_pairs_FDR01_5suppSNP",
                            "num_pairs_FDR01_6suppSNP","num_pairs_FDR01_7suppSNP","num_pairs_FDR01_8suppSNP",
                            "num_pairs_FDR01_9suppSNP","num_pairs_FDR01_10suppSNP")
df_all_FDR01 <- as.data.frame(df_all_FDR01, stringsAsFactors = FALSE)
df_all_FDR01 <- df_all_FDR01 %>% mutate(across(-celltype, as.numeric))

df_long <- df_all_FDR01 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR01_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR01_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.1)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )



rownames(df_all_FDR02) <- NULL
colnames(df_all_FDR02) <- c("celltype","num_pairs_FDR02_3suppSNP","num_pairs_FDR02_4suppSNP","num_pairs_FDR02_5suppSNP",
                            "num_pairs_FDR02_6suppSNP","num_pairs_FDR02_7suppSNP","num_pairs_FDR02_8suppSNP",
                            "num_pairs_FDR02_9suppSNP","num_pairs_FDR02_10suppSNP")
df_all_FDR02 <- as.data.frame(df_all_FDR02, stringsAsFactors = FALSE)
df_all_FDR02 <- df_all_FDR02 %>% mutate(across(-celltype, as.numeric))


df_long <- df_all_FDR02 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR02_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR02_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.2)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )


DT::datatable(df_p0001,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','significant pairs '),options = list(pageLength = 10) )



```


## New EAS traits

```{r fig.width=8, fig.height=5}

df_all_FDR01 <- c()
df_all_FDR02 <- c()
df_p0001 <- c()
for (celltype in celltypes){
  
  df <- readRDS(paste0(folder_ect,celltype,"_ECT_summary_20genes_EAS.RDS"))
  
  tmp_FDR01 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.1,na.rm = T),sum(df$FDR_4suppSNP < 0.1,na.rm = T),sum(df$FDR_5suppSNP < 0.1,na.rm = T),sum(df$FDR_6suppSNP < 0.1,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.1,na.rm = T),sum(df$FDR_8suppSNP < 0.1,na.rm = T),sum(df$FDR_9suppSNP < 0.1,na.rm = T),sum(df$FDR_10suppSNP < 0.1,na.rm = T))
  tmp_FDR02 <- c(unique(df$celltype),sum(df$FDR_3suppSNP < 0.2,na.rm = T),sum(df$FDR_4suppSNP < 0.2,na.rm = T),sum(df$FDR_5suppSNP < 0.2,na.rm = T),sum(df$FDR_6suppSNP < 0.2,na.rm = T),
                 sum(df$FDR_7suppSNP < 0.2,na.rm = T),sum(df$FDR_8suppSNP < 0.2,na.rm = T),sum(df$FDR_9suppSNP < 0.2,na.rm = T),sum(df$FDR_10suppSNP < 0.2,na.rm = T))
  tmp_FDR02[2:length(tmp_FDR02)] <- as.numeric(tmp_FDR02[2:length(tmp_FDR02)]) - as.numeric(tmp_FDR01[2:length(tmp_FDR01)])
  
  df_all_FDR01 <- rbind(df_all_FDR01,tmp_FDR01)
  df_all_FDR02 <- rbind(df_all_FDR02,tmp_FDR02)
  
  df_p0001 <- rbind(df_p0001, df[df$p_ECT < 0.001,])
  
}

df_p0001 <- df_p0001[complete.cases(df_p0001$p_ECT),]

rownames(df_all_FDR01) <- NULL
colnames(df_all_FDR01) <- c("celltype","num_pairs_FDR01_3suppSNP","num_pairs_FDR01_4suppSNP","num_pairs_FDR01_5suppSNP",
                            "num_pairs_FDR01_6suppSNP","num_pairs_FDR01_7suppSNP","num_pairs_FDR01_8suppSNP",
                            "num_pairs_FDR01_9suppSNP","num_pairs_FDR01_10suppSNP")
df_all_FDR01 <- as.data.frame(df_all_FDR01, stringsAsFactors = FALSE)
df_all_FDR01 <- df_all_FDR01 %>% mutate(across(-celltype, as.numeric))

df_long <- df_all_FDR01 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR01_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR01_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.1)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )



rownames(df_all_FDR02) <- NULL
colnames(df_all_FDR02) <- c("celltype","num_pairs_FDR02_3suppSNP","num_pairs_FDR02_4suppSNP","num_pairs_FDR02_5suppSNP",
                            "num_pairs_FDR02_6suppSNP","num_pairs_FDR02_7suppSNP","num_pairs_FDR02_8suppSNP",
                            "num_pairs_FDR02_9suppSNP","num_pairs_FDR02_10suppSNP")
df_all_FDR02 <- as.data.frame(df_all_FDR02, stringsAsFactors = FALSE)
df_all_FDR02 <- df_all_FDR02 %>% mutate(across(-celltype, as.numeric))


df_long <- df_all_FDR02 %>%
  pivot_longer(
    cols = starts_with("num_pairs_FDR02_"),
    names_to = "num_suppSNP",
    values_to = "num_pairs"
  ) %>%
  mutate(
    num_suppSNP = as.numeric(gsub("num_pairs_FDR02_|suppSNP", "", num_suppSNP))
  )


ggplot(df_long, aes(x = celltype, y = num_pairs, fill = factor(num_suppSNP))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Cell type",
    y = "Number of pairs (FDR < 0.2)",
    fill = "Supporting SNPs"
  ) +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )


DT::datatable(df_p0001,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','significant pairs '),options = list(pageLength = 10) )



```


