---
title: "Colocalization analysis"
author: "XSun"
date: "2025-09-26"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

# Introduction

For the pairs with ECT p-value < 0.0005, we performed colocalization analysis on the supporting SNPs. We considered a ±250 kb window around each SNP. Within this region, we conducted association tests for the factor and the SNPs. Finally, we calculated the colocalization posterior probability (PP4) using `coloc` package.

For pairs with ECT p-values greater than 0.0005, we sampled a control set consisting of five times as many pairs as those with ECT p-values below 0.0005 and performed colocalization analysis.


```{r message=FALSE, warning=F}
library(dplyr)
library(ggplot2)

#supp_snp <- readRDS("/project/xinhe/xsun/pathway_factor/analysis/6.coloc/results/coloc_res_supp_snp.RDS")
supp_snp <- readRDS("/project/xinhe/xsun/pathway_factor/analysis/6.coloc_v2/results/coloc_res_supp_snp.RDS")
supp_snp <- supp_snp[order(supp_snp$p_ECT, decreasing = F),]
supp_snp$coloc_PP4 <- round(supp_snp$coloc_PP4, digits = 4)

supp_snp_control <- readRDS("/project/xinhe/xsun/pathway_factor/analysis/6.coloc_v2/results/coloc_res_supp_snp_control.RDS")
supp_snp_control$coloc_PP4 <- round(supp_snp_control$coloc_PP4, digits = 4)

source("/project/xinhe/xsun/pathway_factor/data_v2/traits_finalselection.R")
```


# EUR

## Results for pairs with ECT_p < 0.0005

```{r}

supp_snp_EUR <- supp_snp[supp_snp$trait %in% traits_EUR,]

DT::datatable(supp_snp_EUR,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Colocalization results for the supporting SNPs'),options = list(pageLength = 10) )

# top_snp <- supp_snp_EUR %>%
#   group_by(celltype, trait, factor) %>%
#   slice_max(order_by = coloc_PP4, n = 1, with_ties = FALSE) %>%
#   ungroup()
# 
# ggplot(top_snp, aes(x = coloc_PP4)) +
#   geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white", boundary = 0) +
#   labs(
#     title = "Distribution of Coloc PP4 (top SNP per factor-trait pair)",
#     x = "Coloc PP4",
#     y = "Count"
#   ) +
#   theme_minimal(base_size = 14)

ggplot(supp_snp_EUR, aes(x = coloc_PP4)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white", boundary = 0) +
  labs(
    title = "Distribution of Coloc PP4",
    x = "Coloc PP4",
    y = "Count"
  ) +
  theme_minimal(base_size = 14)
```

## Results for control group

```{r}

supp_snp_control_EUR <- supp_snp_control[supp_snp_control$trait %in% traits_EUR,]

ggplot(supp_snp_control_EUR, aes(x = coloc_PP4)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white", boundary = 0) +
  labs(
    title = "Distribution of Coloc PP4, for control group",
    x = "Coloc PP4",
    y = "Count"
  ) +
  theme_minimal(base_size = 14)
```

## Comparing 2 groups

```{r}

ggplot() +
  geom_density(data = supp_snp_EUR, aes(x = coloc_PP4, color = "ECTp <0.0005 group"), linewidth = 1) +
  geom_density(data = supp_snp_control_EUR, aes(x = coloc_PP4, color = "Control group"), linewidth = 1) +
  scale_color_manual(values = c("ECTp <0.0005 group" = "steelblue", "Control group" = "tomato")) +
  labs(
    title = "Distribution of Coloc PP4",
    x = "Coloc PP4",
    y = "Density",
    color = "Group"
  ) +
  theme_minimal(base_size = 14)


```




# EAS

## Results for pairs with ECT_p < 0.0005

```{r}

supp_snp_EAS <- supp_snp[supp_snp$trait %in% traits_EAS,]

DT::datatable(supp_snp_EAS,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Colocalization results for the supporting SNPs'),options = list(pageLength = 10) )

# top_snp <- supp_snp_EAS %>%
#   group_by(celltype, trait, factor) %>%
#   slice_max(order_by = coloc_PP4, n = 1, with_ties = FALSE) %>%
#   ungroup()
# 
# ggplot(top_snp, aes(x = coloc_PP4)) +
#   geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white", boundary = 0) +
#   labs(
#     title = "Distribution of Coloc PP4 (top SNP per factor-trait pair)",
#     x = "Coloc PP4",
#     y = "Count"
#   ) +
#   theme_minimal(base_size = 14)

ggplot(supp_snp_EAS, aes(x = coloc_PP4)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white", boundary = 0) +
  labs(
    title = "Distribution of Coloc PP4",
    x = "Coloc PP4",
    y = "Count"
  ) +
  theme_minimal(base_size = 14)
```


## Results for control group

```{r}

supp_snp_control_EAS <- supp_snp_control[supp_snp_control$trait %in% traits_EAS,]

ggplot(supp_snp_control_EAS, aes(x = coloc_PP4)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white", boundary = 0) +
  labs(
    title = "Distribution of Coloc PP4, for control group",
    x = "Coloc PP4",
    y = "Count"
  ) +
  theme_minimal(base_size = 14)
```


## Comparing 2 groups

```{r}

ggplot() +
  geom_density(data = supp_snp_EAS, aes(x = coloc_PP4, color = "ECTp <0.0005 group"), linewidth = 1) +
  geom_density(data = supp_snp_control_EAS, aes(x = coloc_PP4, color = "Control group"), linewidth = 1) +
  scale_color_manual(values = c("ECTp <0.0005 group" = "steelblue", "Control group" = "tomato")) +
  labs(
    title = "Distribution of Coloc PP4",
    x = "Coloc PP4",
    y = "Density",
    color = "Group"
  ) +
  theme_minimal(base_size = 14)


```



