---
title: "The number of pairs at ECT FDR < 0.1 & 0.2"
author: "XSun"
date: "2026-01-05"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---


```{r warning=FALSE, message=FALSE}

library(ggplot2)
library(tidyr)

celltypes <- c("B_cell","CD14_positive_monocyte","CD15_positive_leukocyte","platelet","T_cell","thymocyte")
type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")


```



# The number of pairs at ECT FDR < 0.1 & 0.2 for each permutation and real data

We permuted the genotype and expression data in CEDAR by shuffling sample labels 10 times and performed the entire pipeline on each permuted dataset.

We then examined, across the 10 permutations, the number of pairs passing ECT FDR <0.1 & 0.2.


```{r warning=FALSE, message=FALSE}
num_fdr01 <- c()
num_fdr02 <- c()
for (permu in 1:10){
  
  folder_ect_summary <- paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_CEDAR_permu/ECT_summary/perm",permu,"/")
  
  df_all_fdr02 <- c()
  for(celltype in celltypes){
    
    df <- readRDS(paste0(folder_ect_summary, celltype,"_ECT_summary_poolalltraits_selected.RDS"))
    df <- df[complete.cases(df$ECT_FDR_5suppSNP),]
    
    df_fdr02 <- df[df$ECT_FDR_5suppSNP < 0.2,]
    df_all_fdr02 <- rbind(df_all_fdr02,df_fdr02)
    
  }
  
  rownames(df_all_fdr02) <- NULL
  df_all_fdr01 <- df_all_fdr02[df_all_fdr02$ECT_FDR_5suppSNP < 0.1,]
  
  num_fdr01 <- c(num_fdr01, nrow(df_all_fdr01))
  num_fdr02 <- c(num_fdr02, nrow(df_all_fdr02))
}


df <- data.frame(
  Permutation = seq_along(num_fdr01),
  num_fdr01 = num_fdr01,
  num_fdr02 = num_fdr02
)

df_long <- pivot_longer(
  df,
  cols = c(num_fdr01, num_fdr02),
  names_to = "metric",
  values_to = "value"
)

p <- ggplot(df_long, aes(x = Permutation, y = value, color = metric)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  
  # dashed reference lines
  geom_hline(yintercept = 8,
             linetype = "dashed",
             color = "#1f77b4") +
  geom_hline(yintercept = 125,
             linetype = "dashed",
             color = "#d62728") +
  
  # labels on dashed lines
  annotate(
    "text",
    x = Inf, y = 8,
    label = "Original pairs at FDR < 0.1 = 8",
    color = "#1f77b4",
    hjust = 1.05, vjust = -0.5,
    size = 3.5
  ) +
  annotate(
    "text",
    x = Inf, y = 125,
    label = "Original pairs at FDR < 0.2 = 125",
    color = "#d62728",
    hjust = 1.05, vjust = -0.5,
    size = 3.5
  ) +
  
  scale_color_manual(
    values = c(
      num_fdr01 = "#1f77b4",
      num_fdr02 = "#d62728"
    ),
    labels = c(
      num_fdr01 = "FDR < 0.1",
      num_fdr02 = "FDR < 0.2"
    )
  ) +
  
  labs(
    x = "Permutation",
    y = "Number of pairs",
    color = NULL,
    title = "FDR results across permutations"
  ) +
  
  theme_bw(base_size = 12) +
  theme(
    legend.position = "top"
  ) + 
  scale_x_continuous(breaks = 1:10)

p

DT::datatable(df,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;','Number of pairs passing ECT FDR < 0.1 & 0.2'),options = list(pageLength = 10) )

```




# Examing the permutation with large number of pairs

## Permutation 2

```{r warning=FALSE, message=FALSE}

permu <- 2

folder_ect_summary <- paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_CEDAR_permu/ECT_summary/perm",permu,"/")

df_all_fdr02 <- c()
for(celltype in celltypes){
  
  df <- readRDS(paste0(folder_ect_summary, celltype,"_ECT_summary_poolalltraits_selected.RDS"))
  df <- df[complete.cases(df$ECT_FDR_5suppSNP),]
  
  df_fdr02 <- df[df$ECT_FDR_5suppSNP < 0.2,]
  df_all_fdr02 <- rbind(df_all_fdr02,df_fdr02)
  
}

rownames(df_all_fdr02) <- NULL

df_all_fdr02 <- df_all_fdr02[df_all_fdr02$ECT_FDR_5suppSNP < 0.2,]
df_all_fdr02 <- df_all_fdr02[order(df_all_fdr02$ECT_FDR_5suppSNP, decreasing = F),]

DT::datatable(df_all_fdr02,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;',paste0('Pairs with ECT FDR < 0.2 for permutation ', permu)),options = list(pageLength = 10) )
```


## Permutation 7

```{r warning=FALSE, message=FALSE}

permu <- 7

folder_ect_summary <- paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_CEDAR_permu/ECT_summary/perm",permu,"/")

df_all_fdr02 <- c()
for(celltype in celltypes){
  
  df <- readRDS(paste0(folder_ect_summary, celltype,"_ECT_summary_poolalltraits_selected.RDS"))
  df <- df[complete.cases(df$ECT_FDR_5suppSNP),]
  
  df_fdr02 <- df[df$ECT_FDR_5suppSNP < 0.2,]
  df_all_fdr02 <- rbind(df_all_fdr02,df_fdr02)
  
}

rownames(df_all_fdr02) <- NULL

df_all_fdr02 <- df_all_fdr02[df_all_fdr02$ECT_FDR_5suppSNP < 0.2,]
df_all_fdr02 <- df_all_fdr02[order(df_all_fdr02$ECT_FDR_5suppSNP, decreasing = F),]

DT::datatable(df_all_fdr02,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;',paste0('Pairs with ECT FDR < 0.2 for permutation ', permu)),options = list(pageLength = 10) )
```


## Permutation 8

```{r warning=FALSE, message=FALSE}

permu <- 8

folder_ect_summary <- paste0("/project/xinhe/xsun/pathway_factor/analysis/1.ECT_CEDAR_permu/ECT_summary/perm",permu,"/")

df_all_fdr02 <- c()
for(celltype in celltypes){
  
  df <- readRDS(paste0(folder_ect_summary, celltype,"_ECT_summary_poolalltraits_selected.RDS"))
  df <- df[complete.cases(df$ECT_FDR_5suppSNP),]
  
  df_fdr02 <- df[df$ECT_FDR_5suppSNP < 0.2,]
  df_all_fdr02 <- rbind(df_all_fdr02,df_fdr02)
  
}

rownames(df_all_fdr02) <- NULL

df_all_fdr02 <- df_all_fdr02[df_all_fdr02$ECT_FDR_5suppSNP < 0.2,]
df_all_fdr02 <- df_all_fdr02[order(df_all_fdr02$ECT_FDR_5suppSNP, decreasing = F),]

DT::datatable(df_all_fdr02,caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;',paste0('Pairs with ECT FDR < 0.2 for permutation ', permu)),options = list(pageLength = 10) )
```
