---
title: "ACAT results for the selected EUR traits"
author: "XSun"
date: "2025-10-21"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
---

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=15}

suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
suppressMessages(library(tidyr))
suppressMessages(library(forcats))

source("/project/xinhe/xsun/pathway_factor/data_v2/traits_finalselection.R")
#traits <- c(traits_EUR, traits_EAS)
traits <- c(traits_EUR)

celltypes <- c("B_cell","CD14_positive_monocyte","CD15_positive_leukocyte","platelet","T_cell","thymocyte")
type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")
names(type) <- celltypes
names(celltypes) <- type

qqplot <- function(pvalues,title=NULL) {
  
  pval <- pvalues[complete.cases(pvalues)]
  title <- title
  plotdata <- data.frame(observed = -log10(sort(as.numeric(pval))),
                         expected = -log10(ppoints(length(pval))))
  
  qq <- ggplot(plotdata) + theme_bw(base_line_size =0.3) +
    geom_point(aes(expected, observed), shape = 1, size = 1.5,color = "steelblue") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.6, color = "red") +
    ggtitle(title) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = expression(paste("Expected -log"[10],"(p-value)")),
         y = expression(paste("Observed -log"[10],"(p-value)"))) +
    
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12, color = "black") )
  
  return(qq)
  
}

histogram <- function(pvalues,title=NULL,xlab =NULL) {
  
  p <- as.data.frame(pvalues)
  colnames(p) <- "p"
  
  plot <-  ggplot(p, aes(x=p)) + geom_histogram(breaks = seq(0, 1, by = 0.05),color="white",fill = "steelblue")+ 
    theme_bw(base_line_size =0.3) +
    ggtitle(title) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = xlab,
         y = "Count") +
    
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12, color = "black"),
          text= element_text(family="Times")) 
  
  return(plot)
  
}

qqplot_multi <- function(pvalues_list, legend_names = NULL, colors = NULL, title = NULL) {
  
  # Check if legend names are provided, else use default names
  if (is.null(legend_names) || length(legend_names) != length(pvalues_list)) {
    legend_names <- paste("Set", seq_along(pvalues_list))
  }
  
  # Check if colors are provided, else use rainbow colors
  if (is.null(colors) || length(colors) != length(pvalues_list)) {
    colors <- rainbow(length(pvalues_list))
  }
  
  # Initialize an empty list for storing data frames
  plotdata_list <- list()
  
  # Loop through each vector of p-values
  for (i in seq_along(pvalues_list)) {
    pval <- pvalues_list[[i]][complete.cases(pvalues_list[[i]])]
    data <- data.frame(
      observed = -log10(sort(as.numeric(pval))),
      expected = -log10(ppoints(length(pval))),
      set = legend_names[i]  # Use provided legend name
    )
    plotdata_list[[i]] <- data
  }
  
  # Combine all data frames into one
  plotdata <- do.call(rbind, plotdata_list)
  
  # Plotting
  qq <- ggplot(plotdata, aes(x = expected, y = observed, color = set)) + 
    theme_bw(base_line_size = 0.3) +
    geom_point(shape = 1, size = 1.5) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.6, color = "black") +
    ggtitle(title) + 
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = expression(paste("Expected -log"[10], "(p-value)")),
         y = expression(paste("Observed -log"[10], "(p-value)")),
         color = "Groups") +
    scale_color_manual(values = colors) +  # Use provided or default colors
    theme(axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 12, color = "black"),
          axis.title.y = element_text(size = 14),
          axis.text.y = element_text(size = 12, color = "black"))
  
  return(qq)
}

```

For all independent SNPs, we generated 100 random samples using vSampler, matching the number of SNPs in LD and MAF. These randomly sampled SNPs were then used to perform factor–SNP association tests, from which we computed ACAT p-values for each factor–SNP pair. The resulting random ACAT p-values served as the null distributions to correct the original ACAT p-values.

We applied this correction at three levels: pair null, trait null, and global null.

- Under the pair null, the null distribution was derived from the 100 ACAT p-values of each individual factor–SNP pair.
- Under the trait null, all factors corresponding to a given trait were pooled to form the null distribution.
- Under the global null, we pooled all ACAT p-values across factors and traits within each cell type.

When computing ACAT p-values, we used two settings:

- All independent variants associated with the given trait.
- Excluding variants within the MHC region.

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

folder_acat <- "/project/xinhe/xsun/pathway_factor/analysis/2.ACAT_v2/results/acat_correct_EUR/"

acat_allcelltypes <- list()
for (i in 1:length(celltypes)) {
  
  files_acat <- list.files(path = folder_acat, pattern = celltypes[i], full.names = T)
  
  acat_all <- c()
  for(file in files_acat) {
    
    acat <- readRDS(file)
    acat$trait <- sub(".*/[^-]+-(.*)_ACAT_corrected\\.RDS$", "\\1", file)
    acat$celltype <- celltypes[i]
    
    acat_all <- rbind(acat_all,acat)
  }
  
  acat_allcelltypes[[celltypes[i]]] <- acat_all
}

```

# Detailed ACAT p-values for each pair

## B cell (CD19+)

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

celltype <- "B_cell"


acat_celltype <- acat_allcelltypes[[celltype]]
acat_celltype <- acat_celltype[order(acat_celltype$acat_p),]

DT::datatable(acat_celltype[1:100,],caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;',paste0('TOP 100 pairs by Original ACAT p-values -- ', type[celltype])),options = list(pageLength = 10) )


```


## Monocyte (CD14+) 

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

celltype <- "CD14_positive_monocyte"


acat_celltype <- acat_allcelltypes[[celltype]]
acat_celltype <- acat_celltype[order(acat_celltype$acat_p),]

DT::datatable(acat_celltype[1:100,],caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;',paste0('TOP 100 pairs by Original ACAT p-values -- ', type[celltype])),options = list(pageLength = 10) )


```


## Leukocyte (CD15+) 

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

celltype <- "CD15_positive_leukocyte"


acat_celltype <- acat_allcelltypes[[celltype]]
acat_celltype <- acat_celltype[order(acat_celltype$acat_p),]

DT::datatable(acat_celltype[1:100,],caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;',paste0('TOP 100 pairs by Original ACAT p-values -- ', type[celltype])),options = list(pageLength = 10) )


```


## Platelet

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

celltype <- "platelet"


acat_celltype <- acat_allcelltypes[[celltype]]
acat_celltype <- acat_celltype[order(acat_celltype$acat_p),]

DT::datatable(acat_celltype[1:100,],caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;',paste0('TOP 100 pairs by Original ACAT p-values -- ', type[celltype])),options = list(pageLength = 10) )


```

## T cell (CD4+)

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

celltype <- "T_cell"


acat_celltype <- acat_allcelltypes[[celltype]]
acat_celltype <- acat_celltype[order(acat_celltype$acat_p),]

DT::datatable(acat_celltype[1:100,],caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;',paste0('TOP 100 pairs by Original ACAT p-values -- ', type[celltype])),options = list(pageLength = 10) )


```

## T cell (CD8+)

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

celltype <- "thymocyte"


acat_celltype <- acat_allcelltypes[[celltype]]
acat_celltype <- acat_celltype[order(acat_celltype$acat_p),]

DT::datatable(acat_celltype[1:100,],caption = htmltools::tags$caption( style = 'caption-side: left; text-align: left; color:black;  font-size:150% ;',paste0('TOP 100 pairs by Original ACAT p-values -- ', type[celltype])),options = list(pageLength = 10) )


```


# Original ACAT p-values

## MHC included 

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

p1 <- list()
p2 <- list()
for (celltype in celltypes) {
  
  acat_celltype <- acat_allcelltypes[[celltype]]
  
  p <- as.numeric(acat_celltype$acat_p)
  p1[[celltype]] <- histogram(pvalues = p,title = type[celltype],xlab = "ACAT p-values from real data")
  p2[[celltype]] <- qqplot(pvalues = p,title = type[celltype])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```


## MHC removed 

``````{r warning=FALSE, message=FALSE, warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

p1 <- list()
p2 <- list()
for (celltype in celltypes) {
  
  acat_celltype <- acat_allcelltypes[[celltype]]
  
  p <- as.numeric(acat_celltype$acat_p_MHCremoved)
  p1[[celltype]] <- histogram(pvalues = p,title = type[celltype],xlab = "ACAT p-values from real data, MHC removed")
  p2[[celltype]] <- qqplot(pvalues = p,title = type[celltype])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```




# Corrected ACAT p-values (pair null)

## MHC included 

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

p1 <- list()
p2 <- list()
for (celltype in celltypes) {
  
  acat_celltype <- acat_allcelltypes[[celltype]]
  
  p <- as.numeric(acat_celltype$acat_p_correct_pairnull)
  p1[[celltype]] <- histogram(pvalues = p,title = type[celltype],xlab = "Corrected ACAT p-values - Pair Null")
  p2[[celltype]] <- qqplot(pvalues = p,title = type[celltype])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```


## MHC removed 

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=10}

p1 <- list()
p2 <- list()
for (celltype in celltypes) {
  
  acat_celltype <- acat_allcelltypes[[celltype]]
  
  p <- as.numeric(acat_celltype$acat_p_MHCremoved)
  p1[[celltype]] <- histogram(pvalues = p,title = type[celltype],xlab = "Corrected ACAT p-values - Pair Null, MHC removed")
  p2[[celltype]] <- qqplot(pvalues = p,title = type[celltype])
  
}

grid.arrange(grobs=p1, nrow = 2)
grid.arrange(grobs=p2, nrow = 2)

```


# Corrected ACAT p-values (trait null)

## MHC included 

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=15}


for (celltype in celltypes) {
  
  acat_celltype <- acat_allcelltypes[[celltype]]
  
  # traits <- unique(acat_celltype$trait)
  
  p1 <- list()
  p2 <- list()
  for (trait in traits){
    
    acat_trait <- acat_celltype[acat_celltype$trait == trait,]
    p <- as.numeric(acat_trait$acat_p_correct_traitnull)
    p1[[trait]] <- histogram(pvalues = p,title = trait,xlab = "Corrected ACAT p-values - trait Null")
    p2[[trait]] <- qqplot(pvalues = p,title = trait)
  }
  
  grid.arrange(grobs = p1, ncol = 6, top = type[celltype])
  grid.arrange(grobs = p2, ncol = 6, top = type[celltype])
}



```


## MHC removed 

``````{r warning=FALSE, message=FALSE, fig.width=15, fig.height=15}

for (celltype in celltypes) {
  
  acat_celltype <- acat_allcelltypes[[celltype]]
  
  # traits <- unique(acat_celltype$trait)
  
  p1 <- list()
  p2 <- list()
  for (trait in traits){
    
    acat_trait <- acat_celltype[acat_celltype$trait == trait,]
    p <- as.numeric(acat_trait$acat_p_correct_noMHC_traitnull)
    p1[[trait]] <- histogram(pvalues = p,title = trait,xlab = "Corrected ACAT p-values - trait Null, MHC removed")
    p2[[trait]] <- qqplot(pvalues = p,title = trait)
  }
  
  grid.arrange(grobs = p1, ncol = 6, top = type[celltype])
  grid.arrange(grobs = p2, ncol = 6, top = type[celltype])
}


```


